---
title: "Between more than two groups"  # Page title
weight: 203  # Custom weight 
menuTitle: "F-Test" # Short title if the main title is too long
tags: ["mean", "mean comparisons", "anova", "t-Test"]  # Tags describing the content of the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)
library("car")
pss <-  readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

This page presents mean comparisons between more than two groups. This test statistic is called **ANOVA** or **F-Test**. We now want to test the different groups of the `edu` variable.

**Assumptions**:

1. Dependent variable is metric$\checkmark$

2. Independent variable is categorical $\checkmark$

3. Groups are independent of each other $\checkmark$

4. [Normal distribution of the metric variable in each group (only necessary for $n \leq 25$)] $\checkmark$

5. Equality of variance between groups

### Checking the Assumptions
We want to test how the working hours differ between different educational groups. For this, you will use the `edu` variable, which includes multiple educational groups. First, you will perform the Levene's test again. Since you are now including all groups, you can use the formula notation again:
```{r levene, eval=TRUE}
leveneTest(
  pss$wkhtot ~ pss$edu,
  center = "mean"
)
```

{{% expand "How do you interpret the result?" %}}
There are no significant differences between any pair of groups, hence homogeneity of variances is assumed.

$H_0$: All groups have the same variance in the metric variable.
{{% /expand %}}

### Calculating the test (unequal variances)
To calculate the test, you can use the functions `oneway.test()` and `pairwise.t.test()`:
```{r oneway, eval=TRUE}
oneway.test(
  pss$wkhtot ~ pss$edu,
  var.equal = TRUE
)
```
This test indicates (p-value) that there are significant differences between at least two groups (p-value $< 0.05$). To determine between which groups these differences exist, you can use the function `pairwise.t.test()`:
```{r pairwise, eval = TRUE}
pairwise.t.test(
  pss$wkhtot, # metrische Variable zuerst
  pss$edu, # Gruppenvariable als zweites
  p.adj = "holm"  # Korrektur (Standardverwendung)
) 
```
In the test result, you will see a matrix between the groups, where the p-values for the comparison of two groups are entered.

{{% expand "So, between which groups can significant differences be observed?" %}}
Correct, between all groups!
{{% /expand %}}

That's it for this learning block!
