---
title: "One-Sample T-Test"  # Title of the page
weight: 201  # Custom weight
menuTitle: "One-Sample T-Test" # Short title if the title is too long
tags: ["mean", "mean comparisons", "t-Test"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

The one-sample t-test is calculated to test if the empirical value significantly deviates from the true value $\mu$. Therefore, you need knowledge about $\mu$.

Assume that the statistics office states the official average age as $36.8$ years.

What value does the arithmetic mean of age in our dataset take?
The variable for age is `agea`.

{{% tabs groupID=\"q1\" %}}
{{% tab name=\"Task\"%}}
Calculate the arithmetic mean of age from the dataset **PSS**!
{{% /tab %}}
{{% tab name=\"Code\" %}}
\`\`\`R
mean(
  pss$agea,
  na.rm = TRUE
)
\`\`\`
{{% /tab %}}
{{% tab name=\"Solution\" %}}
In the dataset, the average age is $42.83006$ years. So, the value deviates in the dataset.
{{% /tab %}}
{{% /tabs %}}

{{% expand \"Take a moment to think about why the value could deviate in the dataset!\" %}}
Only individuals $\geq 16$ were surveyed in the dataset. Therefore, it is clear that the average age deviates upwards.
{{% /expand %}}

Now you want to test whether this deviation is statistically significant. There are two testing situations for each mean comparison:

- *two-sided*

- *one-sided* (greater or smaller)

You just want to know if the value significantly deviates at first. There is no assumption about the direction. Therefore, you conduct a *two-sided* test.

You can do this with the function `t.test()`:

```{r ttest1, eval=TRUE}
t.test(
  pss$agea, 
  mu = 36.8,
  alternative = "two.sided"
)  
```

Then you will see the **t-value**, the **p-value**, the **confidence interval**, and the **mean**.

The p-value is less than $0.05$, giving you a significant test result. In the output, you see `2.2e-16`. This is nothing but `2.2*10^{-16}`, meaning you need to move the decimal point 16 places to the left. So, it is a value very close to $0$. You can conclude that the mean age in the sample significantly deviates from $\mu$. We have already discussed the possible reasons for this above!

You can also calculate the difference using R:
```{r ttest2, eval=TRUE}
diff_age <- mean(pss$agea, na.rm = TRUE) - 36.8

diff_age
```

The difference is $6.030064$. Since $\mu$ refers to all individuals in Panem while the sample only includes individuals aged $16$ and above, this difference is easily explained.

### Test Alternatives
Alternatively, you can also test *one-sided*:

- if we assume the value is greater than $\mu$, use `greater`

- if we assume the value is smaller than $\mu`, use `less`.

```{r ttest4, eval=TRUE}
# one-sided, greater
t.test(
  pss$agea, 
  mu = 36.8, 
  alternative = "greater"
)

# one-sided, lower
t.test(
  pss$agea,
  mu = 36.8, 
  alternative = "less"
)
```

Let's move on to the t-test for two samples! 
