---
title: "Start: Bar Chart"  # Page Title
weight: 111  # Custom Weight 
menuTitle: "Start" # Short Title if the Title is too long
tags: ["Bar Charts", "ggplot"]  # Tags; Brief description of the content

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Let's start with bar charts. The main function we will consistently use to create a plot with the **library** `ggplot` is `ggplot()`! You can directly output these plots (simple function call), and the image will appear in the **Plots** tab under **Files**. Alternatively, you can also save the plot (like everything else!) in an object.

Saving the plot in an object has several advantages: As you will notice later, a `ggplot` consists of different layers. If you want to make changes or additions and you have saved the previous plot in an object, you only need to call the object and add the modified or new layer. If you don't save your plot in an object, you will have to call the entire code again each time. In this learning module, we work with objects!

Now, let's move on to bar charts. A **bar chart** is a plot of a single categorical variable. You will now create a bar chart of the `edu` variable. 

First, let's take a look at the basic structure of the `ggplot()` function: The function has the data object stored in the first argument (`data = pss`). In the second argument (`mapping`), we specify the axes of the plot. A bar chart has one axis to be determined (x-axis), as the y-axis automatically represents the frequencies in a bar chart. So, you only need to specify the x-axis in the `mapping` here.

```{r basic_structure, eval=FALSE}
# direkter Output
ggplot(
  data = pss,
  mapping = aes(x = edu)
)

```

But when we output this, the following appears:
```{r basic_structure2, eval=TRUE, echo=FALSE}
# direkter Output
ggplot(
  data = pss,
  mapping = aes(x = edu)
)

```

{{% expand \"But why is the plot empty?\" %}}
We have only defined the structure in the lines. Now we need to specify what we want to display in the plot (bars, points, etc.). For **bar charts**, we add the `geom_bar()` function. The individual additional functions and their arguments will be introduced step by step in this learning module.
{{% /expand %}}

So, let's add the instruction that we want a bar chart of the specified data. Now let's also save the whole thing in an object.

```{r barplot1, eval=TRUE}
barplot1 <- ggplot(
  data = pss,
  mapping = aes(x = edu)
) +
  geom_bar()

barplot1
```

We now have a **bar chart** showing the frequency of the values of the variable `edu`. If you prefer to display percentages instead, this is easily achievable. To do this, you need to change the y-axis so that percentages are displayed. Within the `mapping` argument, you also specify the y-axis (because you are changing the default!). Here, you use a shortcut implemented in `ggplot`: `..prop..` automatically displays the proportion.

## Percentages instead of Frequencies
```{r barplot2, eval=TRUE}
barplot2 <- ggplot(
  data = pss, 
  mapping = aes(
    x = edu, 
    y = ..prop.., 
    group = 1
  )
) + 
  geom_bar()

barplot2
```

{{% expand "Within `mapping`, there is also the argument `group = 1`, delete the argument and see what happens" %}}
The argument `group = 1` tells `ggplot` that all values belong to a total group. If you remove the argument, all bars will be of equal height, as the percentage within each value is output, and logically, it is always 100 percent per value.
{{% /expand %}}

We have been talking about **bar charts** all the time, but we have always mistakenly shown **column charts**. We can easily change the columns to bars using the additional function `coord_flip()`:
```{r barplot-coord, eval=TRUE}
 barplot1 + 
   coord_flip()
```

Now let's add some color to the whole thing!
