---
title: "Axes, Legend & Title"   
weight: 114  
menuTitle: "Axes, Legend & Title"  
tags: ["axes", "title", "legend", "bar charts", "ggplot"]   
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,    
  eval = FALSE,  
  message = FALSE,  
  warnings = FALSE  
)

library("tidyverse")
library("beyonce")

pss <- readRDS("./data/pss.rds")

barplotBeyonce <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar() +
  scale_fill_manual(values = beyonce_palette(25))
```


{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

## Axes
Often, the axes need to be adjusted to better represent the data. If you have been experimenting and trying to display values above the bars, you may have noticed that not all values from each bar were displayed. This happened because the y-axis was too short. You can easily limit the axes in the `scale_x_continuous()` function. `continuous` is used here because it deals with the data type (*integer* or *numeric*). To limit the axes, you use the argument `limits = c(...)` and specify the minimum and maximum values.

```{r barplot-axis, eval=TRUE}
barplotBeyonce <- barplotBeyonce +
  geom_text(
    stat = "count", 
    aes(label = ..count..), 
    vjust = -2.5, 
    size = 8,
    color = "white"
  ) +
  scale_y_continuous(
    limits = c(
      0, 
      1750
    )
  )

barplotBeyonce
```

Similarly, you can determine the axis markings in the `scale_y_continuous` function. You do this in the `breaks` argument by simply specifying a sequence.
```{r barplot-ticks, eval=TRUE}
barplotBeyonce <- barplotBeyonce + 
  scale_y_continuous(
    breaks = seq(
      0, 
      1750, 
      100
    ),
    limits = c(
      0,
      1750
    )
  )

barplotBeyonce
```

The same goes for the x-axis: You know that the variable `edu` is a factor with five levels (plus `NA`). So, it is not a continuous variable. Therefore, you use the function `scale_x_discrete()` here.
```{r barplot-axis-discrete, eval=TRUE}
barplotBeyonce <- barplotBeyonce + 
  scale_x_discrete(
    limits = c(
      "ES-ISCED I", 
      "ES-ISCED II",
      "ES-ISCED III",
      "ES-ISCED IV",
      "ES-ISCED V", 
      NA
    )
  ) 

barplotBeyonce
```
So, for non-continuous variables, you need to display the categories that should be shown.

{{% expand \"Think about or try how you can exclude the `NA` category?\" %}}
You may not always want to display the `NA` category. By simply excluding it from the `scale_x_discrete()` function, it will no longer be displayed in the plot.
{{% /expand %}}

{{% expand \"You want to change the order from the highest to the lowest category. How can you do that?\" %}}
Exactly! You just need to change the order within the `scale_x_discrete()` function!
{{% /expand %}}

## Legend
Currently, the legend appears on the right. You can change it as you like using the `legend.position` argument within the additional `theme()` function:
```{r barplot-legend, eval=TRUE}
barplotBeyonce +
  theme(legend.position = "bottom") # left, right, top, or none
```

You can also customize the legend labels:
```{r barlplot-legend2, eval=TRUE}
barplotBeyonce + 
  scale_fill_manual(
    name = "Bildungsniveau",
    labels = c(
      "sehr niedrig", 
      "niedrig",
      "mittel",
      "hoch",
      "sehr hoch", 
      "NA"
    ),
    values = beyonce_palette(25)
  )
```

## Titles 

Now, we just need customized axis labels and titles!
```{r barplot-titles, eval=TRUE}
barplotBeyonce + 
  labs(
    x = "Bildungsniveau", 
    y = "Häufigkeiten",
    title = "My first try with ggplot2"
  )
```

Everything clear? Keep going!

