---
title: "Coloring"  # Page Title
weight: 112  # Custom Weight 
menuTitle: "Coloring" # Short Title if Title is too long
tags: ["coloring", "bar charts", "ggplot"]  # Insert tags here; Short description of what happens on the page
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")
library("RColorBrewer")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

So far, our bar chart is quite dull: the axis labels are unclear, and in gray, it looks dreary. Let's change that now. **It is important** that if we follow the convention in `ggplot`, we can omit explanatory parts like `data = ...` or `mapping = ...`. This means less writing! We simply add color by including the argument `color` within `aes()` and specifying which variable should determine the color.

First, let's add some color:
```{r barplot-clean, eval=TRUE}
barplot <- ggplot(
  pss, 
  aes(
    edu, 
    color = edu
  )
) +
  geom_bar()

barplot
```

We see that the `color` argument only changes the outline. To change the fill, we need to use the `fill` argument:
```{r barplot-fill, eval=TRUE}
barplot <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar() 

barplot
```

Now we have made the graph colorful. However, you can also set the color values yourself if you want to determine how colorful the graph will be. You can do this by using the `fill` argument in the `geom_bar()` function. There, you specify as many colors as you need (no more and no less!). An overview of common color codes can be found here: 

{{%attachments style=\"grey\" title=\"Colors in R\" /%}}

```{r barplot-color, eval=TRUE}
barplot <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar(
    fill = c(
      "steelblue", 
      "darkgoldenrod",
      "seagreen", 
      "red4", 
      "orange",
      "darkslategray2"
    )
  ) 

barplot
```

{{% expand \"But now you want the second bar to be blue. How can you change that?\" %}}
Correct, it's very easy! You just need to change the order!
{{% /expand %}}

You can also store color codes in an object (vector) and then call it in the `ggplot()` function. This keeps the plot code more organized. It is important that the vector must have the exact same length as the number of groups.
```{r barplot-color-vektor, eval=TRUE}
educol = c(
  "steelblue", 
  "darkgoldenrod",
  "seagreen", 
  "red4", 
  "orange",
  "darkslategray2"
)

barplot <- ggplot(
  pss, 
  aes(
    x = edu, 
    fill = edu
  )
) +
  geom_bar(fill = educol) 

barplot
```

When using a color vector, you can also use color vectors that differ from the exact number of categories. If there are too few colors in the color vector, the colors will repeat. However, you cannot use this in the `fill` argument in the `geom_bar()` function; instead, you use the additional function `scale_fill_manual()` for that:
```{r barplot-colorblind, eval=TRUE}
# a colourblind-friendly palettes
cbp1 <- c(
  "#999999", 
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442", 
  "#0072B2",
  "#D55E00",
  "#CC79A7"
)

barplotCb <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) + 
  geom_bar() +
  scale_fill_manual(values = cbp1)

barplotCb
```

When you cannot submit documents digitally, you often use gray graphics to save costs. For this purpose, there is the additional function `scale_fill_grey()`, which automatically outputs the plot in grayscale:
```{r barplot-gray, eval=TRUE}
barplotGray <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) + 
  geom_bar() +
  scale_fill_grey()

barplotGray
```

There are also some **libraries** that contain pre-made color palettes. For example: `RColorBrewer` or [`beyonce`](https://github.com/dill/beyonce). Here, we will first use `RColorBrewer`. `RColorBrewer` is easy to install, while the **library** `beyonce` is installed via a developer tool. Many **libraries** are still in the development stage and are not officially available yet, but they can still be loaded and used through a developer tool.

```{r barplot-install-palettes}
install.packages("RColorBrewer")

# Installation Entwicklertool
install.packages(
  "devtools",
  dependencies = TRUE
)

# Install beyonce palette
devtools::install_github("dill/beyonce")
```

With `RColorBrewer`, you can display the color palettes:
```{r rcolorbrewer, eval=TRUE}
library("RColorBrewer")

display.brewer.all()
```

To use one of these palettes from `RColorBrewer`, you use the function `scale_fill_brewer()` and specify the color palette within it:
```{r barplot-brewer, eval=TRUE}
barplotBrewer <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) + 
  geom_bar() +
  scale_fill_brewer(palette = "Dark2")

barplotBrewer
```

To use the `beyonce` color palette, you proceed differently: First, let's output all color palettes.

```{r beyonce, eval=TRUE}
library("beyonce")

# Ausgabe aller Farbpaletten [beginnend bei 1]
par(mfrow=c(26,5))
for(i in 1:130) print(beyonce_palette(i))
```

In the example, we choose color palette 25. To call the respective color palette, use the `beyonce_palette()` function from the **library** `beyonce`, and as the only argument, provide the number of the color palette. To use it now, use the additional function `scale_fill_manual()` in `ggplot`.

```{r barplot-beyonce, eval=TRUE}
barplotBeyonce <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar() +
  scale_fill_manual(values = beyonce_palette(25)) 

barplotBeyonce

```

Let's go and add value labels! 

{/*examples*/}
