---
title: "Overview & Saving"  # Page Title
weight: 115  # Custom Weight 
menuTitle: "Overview & Saving" # Short Title if the Title is too long
tags: ["saving", "bar charts", "ggplot"]  # Insert tags here; Abbreviation of what happens on the page
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")

pss <- readRDS("./data/pss.rds")

barplotBeyonce <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar() +
  scale_fill_manual(values = beyonce_palette(25))
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Overall, the following steps can be considered:

1. Selection of the variable (grouped or not?)

2. Flipped or not?

3. Limits/order of the y-axis

4. Limits/order of the x-axis

5. Frequencies or percentages?

6. Labels

7. Colors and legend

```{r barplot-final}
barplotFinal <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar() + 
  scale_y_continuous(
    breaks = seq(
      0, 
      1750,
      100
    )
  ) +
  scale_x_discrete(
    limits = c(
      "ES-ISCED I", 
      "ES-ISCED II",
      "ES-ISCED III",
      "ES-ISCED IV",
      "ES-ISCED V"
    )
  ) + 
  scale_fill_manual(
    name = "Bildungsniveau",
    labels = c(
      "sehr niedrig", 
      "niedrig",
      "mittel",
      "hoch",
      "sehr hoch"
    ),
    values = beyonce_palette(25)
  ) +
  geom_text( 
    stat = "count", 
    aes(label= ..count..), 
    vjust = -0.5, 
    size = 3.5,
    color = "darkblue"
  ) +
  labs(
    x = "Bildungsniveau", 
    y = "Häufigkeiten", 
    title = "My first fancy ggplot ohne NA")

barplotFinal
```

## Saving Plots

To save the plot, it is best to use the `ggsave()` function. R saves these files in the *working directory*. If you are unsure where this is, use the `getwd()` function. You can change the *working directory* to any path using `setwd()`. Remember: In RStudio Cloud, the *working directory* is the path of the project you are in! Create a subfolder `img` there!

```{r savepices, eval=FALSE}
getwd()

ggsave("./img/mffggplot.png",  
       width = 8,        
       height = 6,
       units = "in",     
       dpi = 450
       )        
```

As a final step with bar charts, we want to group them by another variable! 
