---
title: "Grouping"  # Title of the page
weight: 116  # Custom weight 
menuTitle: "Grouping" # Short title in case the title is too long
tags: ["grouping", "bar charts", "ggplot"]  # Insert tags here; Abbreviation of what happens on the page
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")

pss <- readRDS("./data/pss.rds")

barplotBeyonce <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar() +
  scale_fill_manual(values = beyonce_palette(25))
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Another grouping variable can easily be added to a plot. This variable should be a factor; if not, the variable will be *coerced* into a factor, which can sometimes lead to issues.
```{r bar-grouped, eval=TRUE}
str(pss$gndr)

barGroup <- ggplot(
  pss, 
  aes(
    edu, 
    fill = gndr
  )
) +
  geom_bar()

barGroup
```

We can also position the sub-bars next to each other instead of stacking them:
```{r bar-group2, eval=TRUE}
barGroup <- ggplot(
  pss, 
  aes(
    edu, 
    fill = gndr
  )
) +
  geom_bar(position = position_dodge())

barGroup
```

Just like before, we can adjust all other options as well:
```{r bar-group-adjustments, eval=FALSE}
barGroup + 
  scale_y_continuous(
    breaks = seq(
      0,
      800, 
      100
    ),
    limits = c(
      0, 
      800
    )
  ) + 
  geom_text(
    stat = "count", 
    aes(label= ..count..), 
    vjust = -1, 
    size = 3.5, 
    position = position_dodge(0.9)
  ) +
  labs(
    x = "Bildungslevel", 
    y = "Häufigkeiten", 
    title = "My first fancy ggplot"
  ) +
  scale_fill_manual(
    name = "Gender", 
    labels = c(
      "Female", 
      "Male"
    ),
    values = beyonce_palette(72)
  )
```

That's how simple it is! Let's move on to the other types of plots!

