---
title: "Value Labels"  # Page Title
weight: 113  # Custom Weight
menuTitle: "Value Labels"  # Short Title if Title is too long
tags: ["value labels", "bar charts", "ggplot"]  # Insert tags here; Abbreviation of what happens on the page
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")

pss <- readRDS("./data/pss.rds")


barplotBeyonce <- ggplot(
  pss, 
  aes(
    edu, 
    fill = edu
  )
) +
  geom_bar() +
  scale_fill_manual(values = beyonce_palette(25)) 
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

You can add individual frequencies per category. To do this, use the `geom_text()` function and simply add it to the last plot: The first argument indicates that the frequencies are used here, and in the second argument, you specify that the output should be the counted frequency (label = ..count..`):

```{r barplot-text, eval=TRUE}
barplotBeyonce +
   geom_text(
     stat = "count", 
     aes(label =  ..count..)
   )
   
```

Now, some of the numbers may not be readable. Let's move the number values and change the font color! We can adjust the position and appearance using various arguments within `geom_text()`:

```{R barplot-text2, eval=TRUE}
barplotBeyonce +
  geom_text(
    stat = "count", 
    aes(label = ..count..), 
    vjust = 1.5, 
    color = "white"
  )

```

{{% tabs %}}
{{% tab name=\"Your Turn Now!\" %}}
Modify the code above so that the numbers are above the bars and the font color is `darkblue`! Try it out a bit before looking at the solution!
{{% /tab %}}
{{% tab name=\"Code\" %}}
\`\`\`R
barplotBeyonce +
  geom_text(
    stat = "count", 
    aes(label = ..count..), 
    vjust = -0.5, 
    color = "darkblue"
  )
\`\`\`
{{% /tab %}}
{{% /tabs %}}

You can adjust the size of the text using the `size` argument. Just change the values in the code to see how it changes!

```{r barplot-text3, eval=TRUE}
barplotBeyonce +
  geom_text(
    stat = "count", 
    aes(label = ..count..), 
    vjust = 1.5, 
    size = 8,
    color = "white"
  )

```

Let's move on to the final settings!
