---
title: "Importing SPSS"  # Title of the page
weight: 124  # Custom weight
menuTitle: "SPSS Import" # Short title if the title is too long
tags: [".sav", "file", "spss", "import"]  # Insert tags here; abbreviation of what happens on the page
---

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

{{%attachments style=\"grey\" title=\"Dataset\" /%}}

The basic functions in R cannot import SPSS files. This means that you now need an additional **library** to have a function that can import the SPSS file. There are several different **libraries** available for this purpose, but here we will focus on the `foreign` **library**.

When you need an additional **library**, you must first install it **once**. This can be done using the `install.packages()` function, where you provide the name of the **library** as an argument - in our case, `foreign`.

```{r install-foreign, eval=FALSE}
install.packages("foreign")
```

{{% expand \"Briefly consider why the name in the argument needs to be enclosed in quotation marks!\" %}}
The name needs to be enclosed in quotation marks because we are passing text into the function. The function then looks for a **library** named `foreign`. If you omit the quotation marks, the function would search for an object named `foreign`, leading to an error message.
{{% /expand %}}

Now you have installed the additional functions from the `foreign` package. To be able to use them, you need to load them briefly every time you want to use them in an R session. This is done with the `library()` function, using the package name as the argument.

```{r library-load, eval=TRUE}
library("foreign")
```

Now we will load the `exam` dataset in R, but this time in SPSS format. The special thing about SPSS is that it often contains descriptions of code values or variable descriptions. However, we do not use these in R; instead, we use the codebook of a dataset for that. You will learn how to work with the codebook in the later examples, but for this simple training dataset, we will skip that for now.

From the `foreign` library, we use the `read.spss()` function, where we also use the `file.path()` function again to be able to use the `path` object. There are two additional arguments in the `read.spss()` function that are new: `use.value.labels` and `to.data.frame`. The latter simply indicates that the final object should be a dataset and not a matrix. With the argument `use.value.labels = FALSE`, we turn off the inclusion of value labels. This way, we avoid issues during import because in SPSS, often only the endpoints of a scale are labeled, and we would end up with mixed variables that would all be read as *character* even though they are numeric.

The import process is quite simple:
```{r eval=FALSE}
examspss <- read.spss(
  "./data/exam.sav", 
  use.value.labels = FALSE,       
  to.data.frame = TRUE
)            
```

That's it for now on data import!
