---
title: "Importing .RData and .rds"  # Page title
weight: 123  # Custom weight
menuTitle: "R-Import" # Short title if necessary
tags: [".RData", "file", ".rds", "import"]  # Tags: brief description of page content

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

exam <- readRDS("../data/exam.rds")   
load("../data/exam.RData")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

{{%attachments style=\"grey\" title=\"Dataset\" /%}}

Once datasets have been saved as R data (`.RData` or `.rds`), importing the data becomes easier. We will now reload the `exam` dataset into the **Environment**.

To do this, we first clean up our **Environment** and remove all objects so that we can see the changes. To do this, we use the `rm()` function (*rm* stands for *remove*). In the function, simply list the objects you want to remove from the Environment, separated by a `,` (comma). Let's delete everything!

```{r clean-env, eval=FALSE}
rm(
  exam, 
  examcsv, 
  examcsv2, 
  path
)
```

When loading R data, it is much easier as the data is directly in a format that R can read. 
When loading a single object (`.rds` file), we use the `readRDS()` function. **Important:** With this function, we need to assign the read data to an object so that it is available in the **Environment**. The advantage is that we can rename the data as we like and are not dependent on the file name. 
```{r import-rds, eval=FALSE}
exam <- readRDS("./data/exam.rds")
```

We could also load an entire environment. This could be, for example, all objects of a session, so what we have just done. This type of data is saved in the `.RData` format and we simply use the `load()` function.

```{r import-rdata, eval-FALSE}
# alternatively entire workspace
load("./data/exam.RData")
```

That's it, this is how we can import data that is available in global formats or in R-specific formats. Finally, we will now learn about importing SPSS datasets, as unfortunately, a number of organizations still provide their datasets in this format. 


