---
title: "Export Tables & Graphics"  # Page Title
weight: 202  # Custom Weight 
menuTitle: "Export" # Short Title if necessary
tags: ["regression", "export", "regression table", "regression plot"]  # Tags: brief description of page content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("modelsummary")
library("dotwhisker")

pss <- readRDS("./data/pss.rds")

olsModel <- lm(
  stfdem ~ 1 + stfeco, #Formel der Regression
  data = pss # Datenobjekt
)  

olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)      
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Often in publications, you not only see regression tables but also plots of individual regression coefficients and their confidence intervals. So far, we have only looked at the tables in the R output. Now you will learn how to get these tables into your text document for your thesis or research paper and how to create a simple plot of the regression coefficients. In the next learning block, we will introduce more advanced plots as well.

## Exporting Models to Text Files
There are various **libraries** that allow you to export results from R regression models into different file formats (e.g., as a *txt* file, as an image file, or as a *tex* file for $\LaTeX$). In this section, we will focus on the `modelsummary` **library**, which makes it easier to export models from R in a visually appealing way to different file formats. If you are interested in alternative **libraries**, here are two options: [`texreg`](https://github.com/leifeld/texreg) and [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) are also available. 

We cannot simply copy and paste tables or results from the console into a text document. However, the **library** `modelsummary` allows for the export of regression tables in various formats that can be used in Word or $LaTeX$.

First, you need to install the **library**. This time, we add a second argument: `dependencies = TRUE`. This indicates that all **libraries** needed by `modelsummary` and not yet installed should be automatically installed.
```{r modelsummary, eval=FALSE}
install.packages(
  "modelsummary",
  dependencies = TRUE
)
library("modelsummary")
```

From the **library**, you use the function `modelsummary()` to display models. Here, we use the results of the regression calculations from the previous chapter, which were the models `olsModel` and `olsModel2`.

To bring this table into an exportable format, simply call the `modelsummary()` function.

```{r ms1, eval=TRUE}
modelsummary(olsModel)
```

You can also easily display multiple models at once. For example, both models simultaneously. Since the models build on each other, it makes sense to present them in a table.

```{r ms2, eval = TRUE}
modelsummary(
  list(
    olsModel, 
    olsModel2
  )
)
```

Alternatively, you can save the list in an object beforehand and then call the object in `modelsummary()`:
```{r ms3, eval=TRUE}
models <- list(
  olsModel, 
  olsModel2
)
```
```{r ms4, eval=FALSE}
modelsummary(models)
```

## Export Formats
In `modelsummary`, you can either directly output a finished file format or the *raw* code for output. In the new (second) argument, specify the file where the output should be saved. This will be automatically saved in the **working directory**.

The available output options are as follows:
```{r ms5, eval=FALSE}
modelsummary(
  models, 
  output = "table.docx" # Word-Datei (package flextable nötig)
)

modelsummary(
  models, 
  output = "table.html" # HTML-Datei
)

modelsummary(
  models, 
  output = "table.tex" # Tex-Datei
)

modelsummary(
  models, 
  output = "table.md" # Markdown-Datei
)

modelsummary(
  models, 
  output = "table.txt" # Text-Datei 
)

modelsummary(
  models, 
  output = "table.png" # Bild-Datei 
)
```

As you can see, there is a format for every need!

If you work with text programs that are based on code (e.g., LaTeX or websites), you can also output the result in the console and then copy it from there. In the second argument, do not specify a file, just a keyword!
```{r ms6, eval=FALSE}
modelsummary(
  models, 
  output = "html"
)

modelsummary(
  models, 
  output = "latex"
)

modelsummary(
  models, 
  output = "markdown"
)
```

## Adjusting Output

### Number of Decimal Places
As a first step, you can adjust the display of values using the `fmt` argument:
```{r adjustment1, eval=TRUE}
# 4 Dezimalstellen nach dem Dezimaltrennzeichen
modelsummary(
  models, 
  fmt = 4
)
```

If there are numbers in the coefficient block above that have a $0$ in the fourth decimal place, they will end after three decimal places. To standardize this, you can enter the following:
```{r adjustment2, eval=TRUE}
# 4 Dezimalzeichen und 0 am Ende
modelsummary(
  models, 
  fmt = "%.4f"
)
```
Alternatively, you can also use scientific notation:
```{r adjustments3, eval=TRUE}
# wissenschaftliche Notation
modelsummary(
  models, 
  fmt = "%.4e"
)
```

If you actually want to switch from the decimal point `.` to `,` in the output, you must change this globally. `modelsummary()` will then adopt this setting. **Important:** You should change it back afterwards!
```{r adjustments4, eval=TRUE}
options(OutDec = ",")

modelsummary(
  olsModel, 
  fmt = 2
)

options(OutDec = ".")
```

### Adjusting Coefficients
By default, each coefficient or value is displayed on a single line. Below that, the standard error is given in round `()` parentheses. Alternatively, you can make individual settings and also display the confidence interval.

To do this, we use two new arguments: `estimates` and `statistics`. In `estimate`, we specify in curly `{...}` braces what should be displayed on the line of each variable. In the example, we display the coefficient and then in square brackets **[]** the lower and upper bounds of the confidence interval. Alternatively, you could also display the p-value (`{p.value}`) or significance stars (`{stars}`). The complete options can be looked up in the [User Guide](https://cran.r-project.org/web/packages/modelsummary/modelsummary.pdf). With the argument `statistics = NULL`, we turn off the previously existing second line after each variable.
```{r adjustments5, eval=TRUE}
modelsummary(
  models,
  fmt = 1,
  estimate  = "{estimate} [{conf.low}, {conf.high}]",
  statistic = NULL
)
```

Now we simply display the regression coefficient plus significance stars:
```{r adjustments6, eval=TRUE}
modelsummary(
  models,
  fmt = 1,
  estimate  = "{estimate}{stars}",
  statistic = NULL
)
```

As you can see, no legend has been added for the significance stars. Whenever you manually add the `{stars}` option, the legend must also be manually added through the `notes` argument:
```{r adjustments7, eval=TRUE}
modelsummary(
  models,
  fmt = 1,
  estimate  = "{estimate}{stars}",
  statistic = NULL,
  notes = list('$^{*} p< 0.1$, $^{**} p<0.05$, $^{***} p<0.01$')
)

```

Alternatively, you can use the `stars` argument, which automatically adds a *Caption*:
```{r adjustments8, eval=TRUE}
modelsummary(
  models, 
  stars = TRUE # standardmäßig an!
) 
```

Lastly, we could also display the confidence interval in a second line. This is done in the `statistics` argument:

```{r adjustments9, eval= TRUE}
modelsummary(
  models,
  fmt = 1,
  estimate  = "{estimate}{stars}",
  statistic = "[{conf.low}, {conf.high}]"
)
```

For the `statistics` argument, the following helpful default outputs are available:

- `statistics = "conf.int"` (confidence interval),
- `statistics = "p.value"` (p-value),
- `statistics = "std.err"` (standard error).

If we use the option `statistics = "conf.int"`, you can additionally specify the interval through the `conf_level` argument:

```{r adjustments10, eval=TRUE}
modelsummary(
  models,
  fmt = 1,
  stars = TRUE,
  statistic = "conf.int", 
  conf_level = .95
)
```

You can also combine different outputs and display multiple lines. To do this, use the `c()` function within the `statistic` argument:

```{r adjustments11, eval= TRUE}
modelsummary(
  models, 
  statistic = c("conf.int",
                "s.e. = {std.error}", 
                "t = {statistic}",
                "p = {p.value}"
  )
)
```

Often, we do not want to display the variable names from the dataset in visualizations, but rather use better strings to help the reader understand the table. For this purpose, we use the `coef_rename` argument:
```{r adjustment12, eval=TRUE}
modelsummary(
  models, 
  coef_rename=c("stfeco" = "Zufriedenheit Ökonomie", 
                "trstlgl" = "Vertrauen Rechtssystem"
  )
)

```

## Dotwhisker
As an alternative to table output, coefficient plots, known as dotwhisker plots, are increasingly used. These plots depict the regression coefficient along with its confidence interval. To create these, you can use the **library** `dotwhisker`. With this **library**, objects from `lm()` functions can be directly plotted.

First, install and load the package:

```{r dotwhisker, eval=FALSE}
install.packages("dotwhisker")

library("dotwhisker")
```

Then, call the `dwplot()` function, which graphically represents the coefficients. This is also a `ggplot`, the functionality of which you will learn in the fifth learning block.

``` {r dwplot, eval=TRUE}
dwplot(olsModel2)
```

You can also directly integrate multiple consecutive models into the plot. To do this, simply use the `list()` function again:
```{r dwplot2, eval=TRUE}
dwplot(
  list(
    olsModel,
    olsModel2
  )
)
```

This way, you can better see if coefficients change when you add additional variables or control variables to the model.

It still looks a bit rough, and you can easily refine it with `ggplot`: Here is an example of how it works, you can see the exact process in Learning Block 5.

We add a line at **0** (significance), change the axis label on the y-axis, adjust the scale on the x-axis, and add a title.
```{r dwplot3, eval=TRUE}
dwplot(
  list(
    olsModel, 
    olsModel2
  )
) +
  # Linie bei 0
  geom_vline(
    xintercept = 0,
    linetype = "dashed"
  ) +
  # Renaming der y-Achse
  scale_y_discrete(labels = rev(
    c(
      "Zufriedenheit \n Ökonomie", 
      "Vertrauen \n Rechtssystem"
    )
  )
  ) + 
  # x-Achse modifiziert
  scale_x_continuous(
    breaks = seq(
      -1, 
      1,
      0.2
    ),
    limits = c(-1, 1)
  ) +
  # Titel und Caption hinzugefügt
  labs(title = "Lin. Regressionsmodelle auf stfdem",
       caption = "Data: Panem Social Survey."
       )
```

This is how you create simple graphics and tables in R that you can easily use in your thesis or assignments. 


