---
title: "Preparations"  # Title of the page
weight: 101  # Custom weight
menuTitle: "Preparations" # Short title if the title is too long
tags: ["preparation", "ggplot"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Before we can start with the graphical representations, we need to make some preparations.

First, load the dataset `pss` into your **Environment**. As a reminder, the dataset is directly available in the RStudio Cloud in the `data` folder or alternatively (if you are working locally) download it from here:

{{%attachments style=\"grey\" title=\"Dataset\" /%}}

`ggplot` is part of the [`tidyverse`](https://www.tidyverse.org) and thus compatible with the other **libraries** from `tidyverse` that we have used so far (`dplyr` and `tidyr`). A detailed description of the **library** and its functionality can be found [here](https://ggplot2.tidyverse.org).

In the learning module, all changes are made step by step, and you can always see them directly in the output of the graph. First, we will focus on four types of graphs, make extensions to them, and in the final chapter, create graphs according to the output of specific methods. 

To start with `ggplot`, you just need to load the **library**. You can do this either through `tidyverse` (recommended) or individually:

```{r library-load, eval = FALSE}
# tidyverse
library("tidyverse")

# singulär
library("ggplot2")
```
