---
title: "tidy up statclass 2"  # Page Title
weight: 324  # Custom Weight 
menuTitle: "" # Short Title if the Title is too long
tags: ["tidyr", "tidyverse", "dataset"]  # Insert tags here; Abbreviation of what happens on the page
---
```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

statclass2 <- readRDS("./data/statclass2.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Where is the problem?
``` {r statclass2a, eval=TRUE}
statclass2
```

> - Variable names are column names!

And here is the solution: Here we also convert to the *long* format again!
```{r statclass2tidy, eval=TRUE}
statclass2Tidy <- statclass2 %>%
  pivot_longer(
    momo:sascha, 
    names_to = "names", 
    values_to = "grade"
  )

statclass2Tidy
```

Are there possibly more problems?

> - `exam` does not contain values, but names of variables, namely `exam1` and `exam2`! Variables that indicate the grade in the exam, whose value is still in `grade`. Therefore, we now use `pivot_wider()` here to make the data *tidy*:

``` {r statclass2tidy2, eval=FALSE}
statclass2Tidy <- statclass2Tidy %>% 
  pivot_wider(
    names_from = exam, 
    values_from = grade
  ) %>% 
  relocate(names) %>% 
  arrange(
    names,
    test
  )

statclass2Tidy
```

Just for practice, you could also transform this back into the original dataset with `pivot_wider()`:

``` {r statclass2tidy3, eval=FALSE}
statclass2re <- statclass2Tidy %>% 
  pivot_wider(
    names_from = test,
    values_from = c(
      exam1, 
      exam2
    )
  )
```
