---
title: "Untidy Datasets"  # Page title
weight: 322  # Custom weight 
menuTitle: "" # Short title if the main title is too long
tags: ["tidyr", "tidyverse", "dataset"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

statclass <- readRDS("./data/statclass.rds")
statclass2 <- readRDS("./data/statclass2.rds")

```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

The following two datasets show data that are not *tidy*. We will clean these up below.
```{r statclass, eval=TRUE}
statclass
```

This dataset is in the so-called *wide* format. That means if we had new exams, we would simply add more columns. However, this can be problematic for processing with R, as we often need a *long* format.

``` {r statclass2, eval=FALSE}
statclass2
```

In this case, we have several issues: Firstly, not all columns contain variables, but observations (`momo`, `kim`, `sascha`), and in `exam`, we find variable names again.

Let's start with `statclass`. 

