---
title: "Tidy Up Statclass"  # Page Title
weight: 323  # Custom Weight 
menuTitle: "" # Short Title if the Title is too long
tags: ["tidyr", "tidyverse", "dataset"]  # Insert Tags here; Abbreviation of what happens on the page
---
```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

statclass <- readRDS("./data/statclass.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

In the table, the grade of each person from each exam can be read. Think briefly about which variables we want to generate from this sentence!

```{r statclassa, eval=TRUE}
statclass
```

> - `names`: momo, sascha, kim
> - `course`: statI, statII, r, spss
> - `grade`: Value depending on the above two.

So there are two pieces of information in the columns `stat1`, `stat2`, `r`, and `spss`. Namely, what type of test it is (implicitly through variable name) and the grade. That is, values are given as variable names here, which violates the rules of a *tidy* dataset. However, in a *tidy* format, we need both pieces of information explicitly! Because the column names here are values (type of exam) and not just names.

To clean this up, we use `pivot_longer()`. First, we specify which columns should be rearranged (in our case `stat1` to `spss`), then in which new variables the names or values should be stored. With `names_to`, we name the new variable that distinguishes the test, and with `values_to`, we name the variable that contains the grades. 

``` {r pivot-longer, eval=TRUE}
statclassTidy <- statclass %>% 
  pivot_longer(
    stat1:spss, 
    names_to = "course", 
    values_to = "grade"
  ) %>% 
  arrange(
    name,
    course
  )

statclassTidy
```

Now we have a *long* format, which often makes data manipulation easier (e.g., with `ggplot2`). **But be careful**: You can't simply calculate the mean of `grade` anymore, as this includes different courses. You have to set conditions when working in the *long* format.

To reverse this process, you can use the `pivot_wider()` function:
``` {r pivot-wider, eval=FALSE}
statclassRe <- statclassTidy %>% 
  pivot_wider(
    names_from = course, 
    values_from = grade, 
  )

statclassRe
```
