---
title: "Descriptive Statistics"  # Title of the page
weight: 201  # Custom weight 
menuTitle: "General" # Short title in case the title is too long
tags: ["descriptive"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")   
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/introduction-to-statistics\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

## Structure and Information of a Dataset
But first, let's apply what we learned earlier and now load the dataset `pss` into the environment! In RStudio Cloud, the dataset is already in your `data` folder. If you are working locally with RStudio, you can find the dataset and the codebook here:

{{%attachments style=\"grey\" title=\"Dataset\" /%}}

{{% tabs groupId=\"tab1\" %}}
{{% tab name=\"Task\" %}}
Load the dataset into the environment!
{{% /tab %}}
{{% tab name=\"Code\" %}}
```R
pss <- readRDS("./data/pss.rds")
```
{{% /tab %}}
{{% /tabs %}}

As a reminder: With the function `head()`, we can get a first look at the data (the first $6$ cases):
```{r eval=FALSE}
head(pss)
```

## Variables and Scale Level

In the **PSS** dataset, there are several variables available. There is a codebook for the dataset, which is stored in RStudio Cloud or can be downloaded via the link above. 
Briefly consider the following variables, what scale level they exhibit, and which data type in R would be suitable for them. 

{{% tabs groupId=\"tab2\" %}}
{{% tab name=\"Variables in the Dataset\" %}}
Variables in the dataset `pss`:

- `wkhtot`

- `gndr`

- `stfdem`

- `trstprl`

{{% /tab %}}
{{% tab name=\"Solution\" %}}
The variables have the following scale levels:

- `wkhtot`: metric

- `gndr`: nominal

- `stfdem`: (pseudo-)metric / ordinal

- `trstprl`: (pseudo-)metric / ordinal

{{% /tab %}}
{{% tab name=\"Data Type in R\" %}}
```

str(pss$wkhtot)

str(pss$gndr)

str(pss$stfdem)

str(pss$trstprl)

```
{{% /tab %}}
{{% /tabs %}}

All variables have the appropriate data type for their scale level.

## Length of a Vector
To display the length of a vector, the `length()` function is used. This function provides the number of values in a vector. In the case of a variable, it indicates the number of observations in that variable.
```{r eval=FALSE}
length(pss$edu)
```

{{% expand \"Can we infer the length of the other vectors, or do we need to display them again?\" %}}
In a dataset or **data frame**, all vectors have the same length (because it's a table!). This means we can infer the length of all vectors from the length of one vector. **However:** this does not mean that all variables have the same number of valid responses, as there may also be `NA's` (missing values).
{{% /expand %}}

Alternatively, you can also determine the length using the `dim()` function. However, the **data frame** object must be called in this case.
```{r dim, eval=TRUE}
dim(pss)
```

The first dimension represents the rows (cases), and the second dimension represents the columns (variables).

Let's continue and display measures of central tendency!
