---
title: "Measures of Dispersion"  # Title of the page
weight: 203  # Custom weight 
menuTitle: "Measures of Dispersion" # Short title in case the title is too long
tags: ["descriptive", "measures of dispersion"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")   
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/introduction-to-statistics\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

### Interquartile Range{.tabset .tabset-fade}
The interquartile range can be calculated using the `IQR()` function:
```{r iqr, eval=FALSE}
IQR(
  pss$stfdem, 
  na.rm = TRUE
)
```

{{% expand \"How do you interpret this number?\" %}}
The distance between the first and third quartiles is 3! This value does not tell us where the first and third quartiles are located.

$$q_{75} - q_{25} = IQR$$
{{% /expand %}}

### Alternative Calculation of IQR
A more customizable alternative is to combine the `diff()` and `quantile()` functions. In the `quantile()` function, you specify the variable ($1^{st}$ argument), the desired quantiles ($2^{nd}$ argument), and finally exclude `NA's`. The function returns two values, and the `diff()` function calculates and outputs the difference between these two values.
```{r iqr-manual, eval=FALSE}
diff(
  quantile(
    pss$stfdem, 
    probs = c(
      0.25,
      0.75
    ),  
    na.rm = TRUE
  )
)   
```

### Standard Deviation & Variance
We obtain the standard deviation and variance using the following two functions:
```{r sd-var, eval=FALSE}
sd(
  pss$stfdem, 
  na.rm = TRUE
)

var(
  pss$stfdem, 
  na.rm = TRUE
)
```

{{% expand "How could you alternatively derive the variance from the standard deviation?" %}}
You know that the standard deviation is just the square root of the variance. So, to derive the variance from the standard deviation, you simply need to square it. This can be implemented as a simple calculation:

```{R}
sd(pss$stfdem, na.rm = TRUE) ^ 2
```
{{% /expand %}}

Now, let's take a step further into data processing and enter the `tidyverse`!
