---
title: "Measures of Central Tendency"  
weight: 202  
menuTitle: "Measures of Central Tendency" 
tags: ["descriptive", "measures of central tendency"]  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")   
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

## Descriptive Statistics in R

Now, we will introduce some functions to display descriptive statistics.

{{% tabs groupId=\"tab1\" %}}
{{% tab name=\"Reminder\" %}}
What measures of central tendency and distribution are there?
{{% /tab %}}
{{% tab name=\"Resolution\" %}}
- Mode, Median, Mean

- Range, Interquartile Range, Quartiles

- Variance, Standard Deviation
{{% /tab %}}
{{% /tabs %}}

### Minimum, Maximum, and Range
The functions for minimum, maximum, and range are:
```{R min-max-range, eval=TRUE}
# Minimum
min(pss$wkhtot)

# Maximum
max(pss$wkhtot)

# Spannweite / Range 
range(pss$wkhtot)
```

Now, try it for the variable `stfdem`. What happens?
```{r nas, eval = TRUE}
min(pss$stfdem)
```
An error message will appear stating that the value cannot be calculated. This is because there are *missing values* (`NA's`) in this variable. Some individuals have not provided a value, which is indicated in R as `NA`. Therefore, a value cannot be calculated.

$\rightarrow$ In order to calculate a value, missing values must be excluded.
```{r na-stfdem2, eval=TRUE}
min(
  pss$stfdem, 
  na.rm = TRUE
)       
## na.rm: NA = missing values, rm = remove      
```

### Mode, Median, and Mean
For the median and mean, there are R base functions that can be used directly:
```{r median-mean, eval=TRUE}
## Median
median(
  pss$stfdem, 
  na.rm = TRUE
) 

## arithm. Mittelwert
mean(
  pss$stfdem, 
  na.rm = TRUE
)
```

For the mode, there is no built-in function, but you can use the `table()` function to display the data table of a variable and then determine the mode(s) from it:
```{r modus, eval=TRUE}
## Mode (no built-in function)
table(pss$stfdem)
max(table(pss$stfdem)) # not helpful for bimodal (etc.) distributions
```

### Quantiles
Sometimes you may want to output different ranges of a variable, such as income in deciles or quintiles. For this purpose, the `quantile()` function is used, combined with the `seq()` function.

The `seq()` function generates a sequence of numbers, which we define using the three arguments *from*, *to*, and *by*. The *from* argument sets the starting value, the *to* argument sets the ending value, and the *by* argument sets the steps. In the example, we go from `0` to `1` in steps of `0.1`!
```{r eval=FALSE}
seq(
  from = 0,
  to = 1, 
  by = 0.1
)

seq(
  by = 0.1, 
  to = 1, 
  from = 0
) 

# wird die Standardreihenfolge (from-to-by) eingehalten, kann man die Argumentbeschriftung auslassen.
seq(
  0, 
  1, 
  0.1
)
```

Now deciles (in `0.1` steps) can be output:
```{r eval= FALSE}
quantile(
  pss$stfdem, 
  probs = seq(
    0,
    1, 
    0.1
  ),    
  na.rm = TRUE
) 
```

### Test
{{% tabs groupId=\"tab2\" %}}
{{% tab name=\"Question\" %}}
What needs to be changed to output quintiles?
{{% /tab %}}
{{% tab name=\"Resolution\" %}}
To achieve quintiles, we go in steps of $0.2$, resulting in five values from $0$ to $1$.

```{r eval=FALSE}
quantile(
  pss$stfdem, 
  probs = seq(
    0,
    1,
    0.2
  ),    
  na.rm = TRUE
)
```
{{% /tab %}}
{{% /tabs %}}

### Summary()
To output a series of descriptive values, you can also use the `summary()` function:
```{r eval=FALSE}
summary(pss$stfdem)
```

Next, let's move on to measures of dispersion!
