---
title: "Count Plots"  # Page Title
weight: 143  # Custom Weight
menuTitle: "Count Plots" # Short Title if necessary
tags: ["count plots", "ggplot"]  # Tags: brief description
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Another great way to represent two pseudometric variables is the **count plot**. Here, combinations that occur more frequently are displayed larger. To do this, we call the `geom_count()` function:

```{r count, eval=TRUE}
countplot <- ggplot( 
  pss, 
  aes(
    trstplt, 
    trstlgl
  )
) + 
  geom_count()

countplot
```

## Coloring
Here we can also add colors:
```{r count2, eval=TRUE}
ggplot(
  pss, 
  aes(
    trstplt, 
    trstlgl)) + 
  geom_count(
    aes(
      color = ..n.., 
      size = ..n..
    )
  ) +  
  guides(color = "legend")  
```

Lastly, you will learn how to represent a bivariate distribution as well!
