---
title: "Grouped Scatterplots"  # Title of the page
weight: 142  # Custom weight
menuTitle: "Groups" # Short title in case the title is too long
tags: ["Groups", "ggplot"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")

pss <- readRDS("./data/pss.rds")

```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

We can also divide **Scatterplots** based on a grouping variable. Here we use the variable of *Districts* from the beginning of the learning module:
```{r scatter-group, eval=TRUE}
groupscatter <- ggplot(
  pss, 
  aes(
    trstplt, 
    trstlgl,
    color = district
  )
) +
  geom_point() +
  geom_jitter(
    width = 1, 
    height = 1
  ) +
  scale_color_manual(
    name = "Distrikte", 
    values = beyonce_palette(25)
  ) +
  scale_x_continuous(
    breaks = seq(
      0, 
      10,
      1
    )
  ) + 
  scale_y_continuous( 
    breaks = seq(
      0,
      10,
      1
    )
  ) + 
  labs(
    x = "Trust in Politicians", 
    y = "Trust in Legal System",
    title = "Group Scatterplot"
  ) 

groupscatter
```
Unfortunately, there is not much to see here as everything is still displayed within one plot. To avoid this, we can use facets!

## Facets
However, we can also display four individual plots in one. To do this, we use the `facet_grid()` function:
```{r scatter-group-facet-grid, eval=TRUE}
groupscatter +
  facet_grid(. ~ district)
```

Another way is possible with the `facet_wrap()` function. This creates evenly spaced fields in both columns and rows, where the individual plots are then displayed:
```{r scatter-group-facet-wrap, eval=TRUE}
groupscatter + 
  facet_wrap(~district)
```

Now we have the legend duplicated, so we want to get rid of one:
```{r scatter-group-facet-grid2, eval=TRUE}
groupscatter2 <- ggplot(
  pss, 
  aes(
    trstplt, 
    trstlgl,
    color = district
  )
) +
  geom_point() +
  geom_jitter(
    width = 1, 
    height = 1
  ) +
  scale_color_manual(values = beyonce_palette(25)) +
  scale_x_continuous(
    breaks = seq(
      0,
      10,
      1
    )
  ) + 
  scale_y_continuous(
    breaks = seq(
      0,
      10,
      1
    )
  ) + 
  labs(
    x = "Trust in Politicians", 
    y = "Trust in Legal System",
    title = "Group Scatterplot"
  ) +
  facet_wrap(~ district) +
  theme(legend.position = "none")

groupscatter2
```

Finally, we can also add a regression line here:
```{r group-scatter-reg2, eval=TRUE}
groupscatter2 + 
  geom_smooth(
    method = "lm", 
    se = TRUE,
    color = "darkgray"
  )
```

Let's move on to another type of visualization for pseudo-metric data!
