---
title: "Bivariate Normal Distribution"  # Page Title
weight: 144  # Custom Weight 
menuTitle: "Bivariate Normal Distribution" # Short Title if the Title is too long
tags: ["bivariate normal distribution", "ggplot"]  # Tags: brief description of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")
pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

In special cases, it is possible to not only output normal distributions of individual variables but also bivariate normal distributions. This allows us to visualize whether the distribution of both variables is also approximately normally distributed or not. To do this, we use the `geom_bin2d()` function:
```{r bi, eval=TRUE}
ggplot(
  pss, 
  aes(
    trstplt, 
    trstlgl
  )
) + 
  geom_bin2d(binwidth = 1) +
  scale_fill_continuous(
    low = "lavenderblush", 
    high = "red"
  )
```

Bivariate normal distributions can also be represented through density using the `stat_density_2d()` function:
```{r bi2, eval=TRUE}
ggplot(
  pss, 
  aes(
    trstplt, 
    trstlgl
  )
) + 
  stat_density_2d(
    aes(fill = ..level..), 
    geom = "polygon"
  ) +
  scale_fill_continuous(
    low = "lavenderblush", 
    high = "red"
  )
```
