---
title: "Start: Scatterplot"  # Page Title
weight: 141  # Custom Weight
menuTitle: "Start" # Short Title if the Title is too long
tags: ["Scatterplots", "ggplot"]  # Insert tags here; Abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

We use **Scatterplots** to represent two (pseudo-)metric variables. To do this, we use the `geom_point()` function.

Often, we only have pseudo-metric variables, but we can still use **Scatterplots** for visualization. We now use `trstplt` and `trstprt`. If you don't remember what these variables stand for, check the codebook!

```{r scatter1, eval=TRUE}
scatter <- ggplot(
  pss, 
  aes(
    trstplt, 
    trstprt
  )
) + 
  geom_point()

scatter
```

## Jitter
To better identify data points on the plot, we need to scatter the data points so they do not overlap. Since pseudo-metric variables usually have only integer values, data pairs may overlap. We use the `geom_jitter()` function for this:

```{r scatter2, eval=TRUE}
scatter <- scatter +
  geom_jitter(
    width = 0.3,
    height = 0.3
  )

scatter
```

In the arguments of `geom_jitter()`, specify how far you want the data points to jitter. Just try a few times with different values. 

{ /* examples */ }

## Labels
Now we add labels and titles.
```{r scatter3, eval=FALSE}
scatter +
  geom_point() +
  geom_jitter(
              width = 0.3, 
              height = 0.3
              ) +
  labs(
       x = "Trust in Politicians", 
       y = "Trust in Legal System", 
       title = "Trust Scatterplot"
       )
```

And we change the appearance of the title: Within the `theme()` function, we modify the display. You will learn more about what the arguments do in Chapter 3!

```{r scatter4, eval=TRUE}
scatter <- scatter +
  labs(
    x = "Trust in Politicians",
    y = "Trust in Legal System", 
    title = "Trust Scatterplot"
  ) +
  theme(
    plot.title = element_text(
      face = "bold", 
      hjust = 0.5, 
      size = 16
    )
  )
scatter
```

We also specify the data source. We do this using the `lab()` function and the `caption` argument:
```{r scatter5, eval=TRUE}
scatter <- scatter + 
  labs(caption = "Data source: Panem Social Survey.")

scatter
```

## Axes
The variable has only integer values, but the markings are always at the midpoint. Let's change that now:
```{r scatter6, eval=TRUE}
scatter <- scatter +
  scale_y_continuous(
    breaks = seq(
      0, 
      10, 
      1
    )
  ) + 
  scale_x_continuous(
    breaks = seq(
      0,
      10,
      1
    )
  )

scatter
```

## Regression Line
We can also already plot a regression line for the relationship between the two variables. To do this, we use the `geom_smooth()` function. In the `method` argument, we specify that it is a linear model (`lm`), the confidence interval should be plotted (`se = TRUE`), and we set colors.
```{r scatter7, eval=TRUE}
scatter + 
  geom_smooth(
    method = "lm", 
    se = TRUE, 
    color = "darkred",
    fill = "orange"
  )
```

Let's continue, and now you add a grouping variable!
