---
title: "Cross Tables"  # Title of the page
weight: 102  # Custom weight
menuTitle: "Cross Tables" # Short title if the title is too long
tags: ["cross tables"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

xfun::pkg_load2(
                 c("htmltools", "mime")
               )
library("xaringanExtra")
library("gmodels")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

A cross table provides information about a bivariate distribution, that is, the distribution of two variables. To obtain a cross table, we just need to add a second variable in the `table()` function: Here, we display a cross table between satisfaction with democracy and gender:

```{r crosstab1,eval=TRUE}
table(
  pss$stfdem, 
  pss$gndr
)
```

{{% expand \"Which variable is the column variable?\" %}}
The second variable in the function is the column variable (here `gndr`). The convention in creating cross tables requires that the independent variable goes into the column. Therefore, we assume that gender has an effect on how satisfied we are with democracy.
{{% /expand %}}

As with everything in R, we can also save this table as an object:
```{r crosstab2, eval=TRUE}
mytable <- table(
  pss$stfdem, 
  pss$gndr
)
```

To display percentages instead of frequencies, the `prop.table()` function is used. This function can output column, row, or total percentages. We can either call a table object or use the `table()` function within the function. Here, we call the row percentage by specifying the value $1$ in the last argument.
```{r crosstab3, eval=TRUE}
# Zeilenprozentuierung
prop.table(
  mytable, 
  1
) 
```
```{r crosstab4, eval=FALSE}
# alternativer Weg mit table()-Funktion
prop.table(
  table(
    pss$stfdem, 
    pss$gndr
  ), 
  1
)
```

The column percentages can be displayed by changing the second argument to the value $2$:
```{r crosstab5, eval=TRUE}
prop.table(
  mytable,
  2
)
```

If no second argument is provided, the total percentages will be automatically displayed:
```{r crosstab6, eval=TRUE}
prop.table(mytable)
```

## Marginal Frequencies
The marginal frequencies can be displayed using the `margin.table()` function. Again, the value $1$ in the second argument represents row marginal frequencies, and the value $2$ displays the column marginal frequencies:
```{r crosstab7, eval=TRUE}
# Zeilenrandhäufigkeiten
margin.table(
  mytable, 
  1
)

# Spaltenrandhäufigkeiten
margin.table(
  mytable, 
  2
)
```

## Other Libraries
There are various **libraries** that allow for a more aesthetically pleasing representation of relative frequencies. One of them is the *library* `gmodels`:
```{r gmodels, eval=FALSE}
if(!require("gmodels")) install.packages("gmodels")
library("gmodels")
```

From the **library**, we use the `CrossTable()` function. This function outputs all the information mentioned above at once:
```{r gmodels2, eval=TRUE}
CrossTable(
  pss$stfdem, 
  pss$gndr
)
```
