---
title: "Measures of Association"  # Page title
weight: 103  # Custom weight 
menuTitle: "Measures of Association" # Short title if the main title is too long
tags: ["measures of association", "chi-square", "cramer's v"]  # Tags: brief description of page content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

xfun::pkg_load2(
                 c("htmltools", "mime")
               )
library("xaringanExtra")
library("DescTools")

pss <- readRDS("./data/pss.rds")

mytable <- table(
  pss$stfdem, 
  pss$gndr
)
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

With measures of association, you can test whether and to what extent two variables are related. In your statistics lecture, you learned about $\chi^2$ as a first measure to test for an association. Another measure you are familiar with for the strength of an association is Cramer's V.

## $\chi^2$ Independence Test
Cross-tabulations can display relationships between two variables. This relationship can also be statistically tested using the $\chi^2$ independence test.

The null hypothesis of the $\chi^2$ independence test is:

- $H_0:$ Variables are statistically independent.

We can perform the test using the `chisq.test()` function and directly use the cross-tabulation object:
```{r chi1, eval=TRUE}
chi1 <- chisq.test(mytable)

chi1
```

Alternatively, you can also specify the two variables:
```{r chi2, eval=FALSE}
chi2 <- chisq.test(
  pss$stfdem, 
  pss$gndr
)

chi2
```

{{% expand \"How do we interpret this test?\" %}}
First, look at the p-value. In this example, $p = 0.1674$, which is greater than $0.05$. This means that there is no significant test value, and the null hypothesis can be retained.

From this, we can conclude that the variables are statistically independent.
{{% /expand %}}

## Assumptions of the $\chi^2$ Independence Test
The calculation of the $\chi^2$ independence test requires two assumptions:

1. At least $10$ observations in each cell.

$\Rightarrow$ Cross table! ($\checkmark$)

2. At least $5$ expected observations in each cell.

The function `chisq.test()` stores various values. Therefore, an associated object has the data type **list**. This data type is new to you. It is another data type that plays a significant role in R. In the **list** data type, multiple pieces of information can be stored, similar to a list on paper. You can use `str()` to display what has been stored in the list:
```{r list1, eval=TRUE}
str(chi1)
```

The **list** data type can store different elements as a list, making these elements individually accessible. For example, different vectors can be stored in a list. The individual parts of a list can be output using the `ls()` function and addressed with the `$` sign.

```{r list2, eval=TRUE}
ls(chi1)
```

To see the expected values, we access `expected` from the object `chi1`.
```{r list3, eval=TRUE}
chi1$expected
```

You can see that the condition of having at least $5$ cases in each cell of the expected values is also met.

## Cramer's V
So far, we have only conducted the $\chi^2$ independence test. However, in addition to the general relationship, it is often of interest to know how strong this relationship is or in which direction it goes.

We calculate **Cramer's V** for the strength of the relationship between two at least nominally scaled variables. We use the function `CramerV()` from the **library** `DescTools`. You specify the two variables and in the third argument, you set the confidence interval (in this case, $0.95`):

```{r eval=FALSE}
install.packages("DescTools")
library("DescTools")
```
```{r cramer1, eval=TRUE}
CramerV(
  pss$stfdem,       
  pss$gndr,          
  conf.level = 0.95 # Konfidenzintervall
)  
```

We interpret the test as follows: The first column shows the value for **Cramer's V**, and the second and third columns show the confidence interval for this value. If the interval does not include the value $0$, the result is significant. In this case, we see that the result is not significant.

The following boundaries apply for interpreting the value:

| Lower Bound | Upper Bound | Interpretation    |
| :---------: | :---------: | ----------------- |
| $0$           | $0.1$         | no association    |
| $0.1$         | $0.3$         | weak              |
| $0.3$         | $0.6$         | moderate           |
| $0.6$         | $1$           | strong            |

If not only the strength of the association should be interpreted but also the direction of the relationship, a correlation must be calculated. 