---
title: "Tables"  # Title of the page
weight: 101  # Custom weight
menuTitle: "Tables" # Short title if the title is too long
tags: ["tables"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

xfun::pkg_load2(
                 c("htmltools", "mime")
               )
library("xaringanExtra")
library("summarytools")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Before you start calculating measures of association, you will first create tables. This is sometimes relevant for graphical representation. Often, tables displaying frequencies of variables are chosen for representation. For a simple (frequency) table, we call the `table()` function:
```{r table1, eval=TRUE}
table(pss$stfdem)
```
The first row contains the code values, and the second row contains the frequencies. Listed here are the valid cases, i.e., each value that is not set to *NA*.

How many valid cases do we have? How many NA's?

To test this, we use the `sum()` and `length()` functions:
```{r table2, eval=TRUE}
# gültige Fälle aus der Tabelle
sum(
  table(
    pss$stfdem
  )
) 

# Summe der NA's
sum(
  is.na(
    pss$stfdem
  )
)

# Gesamtlänge: Gültige Fälle + NA's
length(pss$stfdem)
```

Alternatively, you can expand the `table()` function with the argument `useNA = "ifany"`:
```{r table3, eval=TRUE}
table(
  pss$stfdem,
  useNA = "ifany"
)
```

## Table with **library** `summarytools`
To obtain a structured output, you can use the **library** `summarytools`. You will get a view similar to SPSS. The **library** must be installed or loaded first:
```{r eval=FALSE}
install.packages("summarytools")
library("summarytools")
```

Then the `freq()` function is used:
```{r st-freq, eval=TRUE}
freq(pss$stfdem)
```

The first column contains the code values. The second column shows the frequencies, the third column shows the percentages (of valid cases), and the fourth column shows the cumulative percentages (of valid cases). Columns $5$ and $6$ show the percentages and cumulative percentages of all cases (including `NA's`).

In the next step, you will learn about cross-tabulations!
