---
title: "Multivariate lin. Regression with dichotomous categorical variables"  # Page title
weight: 104  # Custom weight 
menuTitle: "Dichotomous Categorical Variables" # Short title if the title is too long
tags: ["regression", "categorical", "dichotomous"]  # Tags; brief description of the content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")

olsModel <- lm(
  stfdem ~ 1 + stfeco, #Formel der Regression
  data = pss # Datenobjekt
)  

olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)      
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

In social sciences, we often use variables with nominal or ordinal scale levels. We can also add these to the **linear regression model** as independent variables. We just need to understand the additional effect we are calculating. We need to create *dummy* variables that indicate the difference of one attribute of the variable in reference to another attribute (reference category). This is easy to implement in the `lm()` function.

We want to include the variable `gndr` in our model (*female*/*male*). What theoretical assumption can we make about the effect?

{{% expand "What do we equalize in the regression equation?" %}}

$$Y = \beta_0 + \beta_1*X_1 + \beta_2*X_2 + \beta_3*X_3+ e$$

$$stfdem = \beta_0 + \beta_1*stfeco + \beta_2*trstlgl + \beta_3*gndr + e$$
{{% /expand %}}

As we already know, a dichotomous variable does not have a linear relationship with a metric variable. Therefore, we need **dummy variables**. We calculate a model where the additional effect of one category compared to the other category is estimated. This effect is a constant effect!.

The variable `gndr` has the following categories:

- *female*

- *male*

## Calculating the Model
Here, we only need to add the variable `gndr` according to the extended equation:
```{r olsModel3, eval=TRUE}
olsModel3 <- lm(
  stfdem ~ 1 + stfeco + trstlgl + gndr,   
  data = pss
)            

summary(olsModel3)
```

{{% tabs %}}
{{% tab name="Question" %}}
What is the reference category?

What effect are we calculating with the variable `gndrmale`?

{{% /tab %}}
{{% tab name="Solution" %}}
The reference category is *female*, meaning female respondents.

The additional effect is calculated for individuals who are male (compared to female individuals). Therefore, male individuals have a satisfaction level $-0.08020$ lower than female individuals.
{{% /tab %}}
{{% tab name="Interpreting the Model" %}}
How do we interpret the model?

Write a few lines in the script!
{{% /tab %}}
{{% tab name="Sample Solution" %}}

The model explains $45.98 %$ of the variance in the variable `stfdem`. Satisfaction with economic performance (`stfeco`) and trust in the legal system (`trstlgl`) have a significant effect on satisfaction with democracy (`stfdem`). The effect of `stfeco` is positive ($\beta_1 = 0.87435$), the effect of trust in the legal system ($\beta_2 = -0.04137$), and the effect for male individuals ($\beta_3 = -0.08020$) are both negative. Individuals with higher trust or who are male have slightly lower satisfaction.
{{% /tab %}}
{{% /tabs %}}

{{% expand "If you compare the formula above with the task again, something should catch your attention! What needs to be changed in the formula to actually fit the regression model?" %}}
In the equation above, the variable gender (`gndr`) is listed. However, we actually add a variable `gndrmale`, which indicates in category `1` that the person is male and in category `0` that the person is not male (so in this dichotomous example, *female*).

The formula would therefore be correct as follows:
$$stfdem = \beta_0 + \beta_1*stfeco + \beta_2*trstlgl + \beta_3*male + e$$
{{% /expand %}}

Why this is relevant, you will see on the next page! Now you have learned how to include dichotomous categorical variables in the regression model, but how does it work with polytomous categorical variables?
