---
title: "Multivariate Linear Regression"  # Page title
weight: 103  # Custom weight 
menuTitle: "Multivariate Example" # Short title if necessary
tags: ["regression", "multivariate"]  # Tags: brief description of page content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}}  

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

In social sciences, we usually do not calculate bivariate models since cause-effect relationships in social phenomena are never bivariate. The previous model was just for easier access.

Now, we expand the model and calculate a **multivariate linear regression**. We want to include the variable `trstlgl` in the model. What effect do we theoretically expect from the variable `trstlgl`?
```{r eval=FALSE}
cor(
  pss$trstlgl, 
  pss$stfdem,
  method = "pearson", 
  use = "complete.obs"
)
```

How do we interpret the result?

$\Rightarrow$ The correlation value between `trstlgl` and `stfdem` indicates a negative correlation, but it is close to $0$, suggesting that there is no relationship between the two variables.

### Model Expansion
We expand the model by adding the variable `trstlgl`.

{{% expand \"What does our linear equation look like?\" %}}

$$Y = \beta_0 + \beta_1*X_1 + \beta_2*X_2 + e$$

$$stfdem = \beta_0 + \beta_1*stfeco + \beta_2*trstlgl + e$$
{{% /expand %}}

We simply implement this in the `lm()` function:
```{r linreg1, eval=TRUE}
olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)            

summary(olsModel2)
```

{{% tabs %}}
{{% tab name=\"Task\" %}}
How do we interpret the result? Write a few lines in your script!

{{% /tab %}}
{{% tab name=\"Solution\" %}}
The model can explain $45.96 %$ of the variance in `stfdem`. With each increase in `stfeco` (satisfaction with economic performance), `stfdem` increases by $0.87361$ points. With each increase in trust in the legal system (`trstlgl`), satisfaction with democracy decreases by $-0.04212$. Both effects are significant ($p<0.05$).
{{% /tab %}}
{{% /tabs %}}

So, you can now also calculate multivariate models and know how to interpret them!
