---
title: "Multivariate lin. Regression with polytomous categorical variables"  # Page title
weight: 105  # Custom weight
menuTitle: "Polychotomous Categorical Variables" # Short title if the title is too long
tags: ["regression", "categorical", "polytomous"]  # Tags: brief description of the content
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")

olsModel <- lm(
  stfdem ~ 1 + stfeco, #Formel der Regression
  data = pss # Datenobjekt
)  

olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)      

olsModel3 <- lm(
  stfdem ~ 1 + stfeco + trstlgl + gndr,   
  data = pss
)           
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

We can add not only dichotomous but also polytomous categorical variables to the regression model. Now, we want to include the variable `edu` in the model. This variable represents the highest level of education attained by the surveyed individual. What theoretical assumptions could we make about the effect of `edu`?

## Calculating the Model
We simply add the variable to the `lm()` function as before:
```{r olsModel4, eval=TRUE}
olsModel4 <- lm(
    stfdem ~ 1 + stfeco + trstlgl + gndr + edu,   
    data = pss
)
summary(olsModel4)
```

{{% tabs %}}
{{% tab name=\"Question\" %}}
How did the `lm()` function incorporate the polytomous variable into the model?

What is the reference category?

And what would the regression equation look like?
{{% /tab %}}
{{% tab name=\"Solution\" %}}
It calculates the additional effect of the four highest categories (`ES-ISCED II`, `ES-ISCED III`, `ES-ISCED IV`, and `ES-ISCED V`) compared to the lowest category (`ES-ISCED I`). 

The reference category is therefore `ES-ISCED I` (omitted category).

$$\begin{align*}stfdem = &\beta_0 + \beta_1*stfeco + \beta_2*trstlgl + \beta_3*gndr + \\ &\beta_4*eduLevelII + \beta_5*eduLevelIII + \\ &\beta_6*eduLevelIV + \beta_7*eduLevelV + \\&e \end{align*}$$
{{% /tab %}}
{{% /tabs %}}

## Changing the Reference Category
In this example, the lowest (or first) category was automatically chosen as the reference category. But what if you want, for example, the middle category (`ES-ISCED III`) as the reference? You can easily do this using the `relevel()` function. In the first argument, specify the data source (variable `pss$edu`), and in the second argument `ref`, specify the corresponding category (i.e., `"ES-ISCED III"`). **Important:** You must save this with the assignment arrow in the variable in the dataset!

```{r relevel, eval=TRUE}
pss$edu <- relevel(
  pss$edu, 
  ref = "ES-ISCED III"
) 
```

Then you just need to recalculate the model:
```{r olsModel5, eval=TRUE}
olsModel5 <- lm(
  stfdem ~ 1 + stfeco + trstlgl + gndr + edu,
  data = pss
)

summary(olsModel5)
```

{{% tabs %}}
{{% tab name=\"Interpretation\" %}}
Interpret the re-specified model. Write a few sentences in the script.

{{% /tab %}}
{{% tab name=\"Sample Solution\" %}}

The model explains $46.35 %$ of the variance in `stfdem`. The effect of `stfeco` is positive and significant ($p<0.001$). Individuals with higher trust in the legal system have lower satisfaction ($\beta_2 = -0.044393$, $p<0.01$). Male and female respondents do not have different satisfaction levels ($\beta_3 = 0.001836$. $p>0.05$). Compared to individuals with a medium level of education, individuals with very low education level (`ES-ISCED I`) and low education level (`ES-ISCED II`) have lower trust ($\beta_4 = -0.343037$ and $\beta_5 = -0.174643$ respectively). Both effects are significant. Individuals with the highest education level have significantly higher trust than those with a medium education level ($\beta_7 = 0.527465$, $p<0.001$). Individuals with the second-highest education level have slightly higher trust ($\beta_6 = 0.076024$), but this effect is not significant.
{{% /tab %}}
{{% /tabs %}}

So now you can also add polytomous categorical variables and interpret the regression model in the R output!
