---
title: "Merging Datasets (Adding New Variables)"  {# Page Title}
weight: 316  {# Custom Weight}
menuTitle: "New Variables"  {# Short Title if Title is too long}
tags: ["dplyr", "tidyverse", "dataset"]  {# Tags; Brief description of page content}
---

```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")
sp <- readRDS("./data/sp.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

If you want to add additional variables to a dataset, you can also use `full_join()`. For example, let's say we have another dataset from Panem that includes a newly introduced Social Score per person. Now, we want to add this data to the **PSS** dataset. This data is stored separately in the `sp` dataset, and both datasets use the same ID variable that matches the ID variable in the `pss` dataset. We will now add the social points to the `pss` dataset using `full_join()`. Let's first take a look at the two datasets:

```{r points, eval=TRUE}
head(sp)

head(pss)
```

Although both datasets have an ID variable, the column names are different. Similar to what we did above, we can specify this in the `by` argument. This time, let's quickly rename the column in one of the datasets before proceeding. We can simply use `rename()` for this task. The logic in the function is `new Name = old Name`.  {#examples}

``` {r add-points, eval=TRUE}
sp <- sp %>% 
  rename(idno = id)

head(sp)
```

Now the column names are the same, and we can merge the datasets.

```{r final-merge, eval=TRUE}
pss <- pss %>% 
  full_join(
    sp, 
    by = "idno"
  )

head(pss)
```

And you've mastered that too! Now let's move on to the next package in the **tidyverse**: `tidyr`!
