---
title: "Most Common Functions"  # Page title
weight: 311  # Custom weight 
menuTitle: "" # Short title if necessary
tags: ["dplyr", "tidyverse", "functions"]  # Tags describing the content of the page
---
```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
library("rmarkdown")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

x <- c("select()", "slice()", "filter()", "arrange()", "mutate()", "summarize()", "group_by()")
y <- c("Spalte(n) wählen", "Zeile(n) wählen", "Zeile(n) filtern", "Zeile(n) ordnen", "neue Spalten / Variablen", "Werte zusammenfassen", "gruppieren")
data <- data.frame(x, y)
colnames(data) <- c("Funktion", "Operation")
rm(x, y)

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

## Prepare Data

In most cases, preparatory work on the dataset is necessary to perform analysis steps. These steps are often of great importance and can be quite time-consuming.

There are several packages that allow for data manipulation. One of the most well-known and commonly used packages due to its clarity is `dplyr`. The idea behind `dplyr` is that it processes only tables (**data frames** or **tibbles**) and facilitates processing through what is known as **piping**. The code in `dplyr` is structured into a long chain, which may initially seem unnecessary. However, this makes the code easy to read, and only a few functions need to be learned, which together can cover the most common processing steps. By breaking down the process into steps, each function has a well-defined scope that is easy to understand. By combining multiple functions, more complex steps can also be performed. Additionally, functions from the `dplyr` package always return a **tibble** (table).

The extensive documentation for `dplyr` can be found [here](https://dplyr.tidyverse.org/).

## Most Common Functions

The most common functions used with `dplyr` are the following:
```{r relevant-functions, eval=TRUE, echo=FALSE}
kable(data)
```

These will now be introduced to you step by step before we learn about **piping**, the second principle of `dplyr`. These examples are somewhat artificial, as `dplyr` is almost exclusively used with **piping**. However, for understanding the functions, it is helpful to get to know them first.

In all examples, we use a fictional dataset that surveyed the study motivation (`mot`) of students at the universities of Gießen, Marburg, and Frankfurt.

First, let's get an overview of the dataset here:
```{r dataframe, echo=FALSE, eval=TRUE}
paged_table(
  pss,
  options = list(rows.print = 15)
)
```

The dataset includes questions about satisfaction with democracy (`stfdem`), district (`district`), trust in parliament (`trstprl`), and gender (`gndr`), among others.

The *package* `dplyr` can be loaded individually, but it is recommended to load the *package* `tidyverse`, which also loads additional packages like `ggplot2` directly.
```{r eval=FALSE}
# install.packages(
#   "tidyverse", 
#   dependencies = TRUE
# )
library("tidyverse")
```

### select()
With the `select()` function, you can choose one or more columns from a dataset. The output is always a **tibble**.

In this example, we only want to display the variables gender (`gndr`) and district (`district`).

```{r head, eval=TRUE}
head(pss)
```
```{r select, eval=FALSE, echo=TRUE}
select(
  pss,
  c(
    gndr, 
    district
  )
)
```
```{r select-table, eval=TRUE, echo=FALSE}
paged_table(
  select(
    pss, 
    c(
      gndr, 
      district
    )
  ),
  options = list(rows.print = 15)
)
```

### slice()
On the other hand, with `slice()`, you can display individual rows. For example, rows $50$ to $55$, or in combination with the `seq()` function, every $100^{th}$ row.

```{r slice, eval = TRUE}
slice(
  pss,
  50:55
)
```

```{r slice2, eval = TRUE}
slice(
  pss, 
  seq(
    0, 
    1000,
    100
  )
)
```

### filter()
In addition to selecting specific rows (cases) or columns (variables), we can narrow down the dataset with `filter()`. For example, we can display only the cases that live in district 1. 

```{r filter, eval=FALSE, echo=TRUE}
filter(
  pss, 
  district == "Distrikt 1"
)
```
```{r filter-table, eval=TRUE, echo=FALSE}
paged_table(
  filter(
    pss,
    district == "Distrikt 1"
  ),
  options = list(rows.print = 15)
)
```

We can also introduce multiple conditions. For example, only people living in District 5 and are male (*male*).

```{r filter2, eval = FALSE, echo=TRUE}
filter(
  pss,
  district == "Distrikt 5" & gndr == "male"
)
```
```{r filter2-table, eval=TRUE, echo=FALSE}
paged_table(
  filter(
    pss,
    district == "Distrikt 5" & gndr == "male"
  ),
  options = list(rows.print = 15)
)
```

All known logical connections work here as well. As a reminder, here are the logical connectors:

* logical *and*: `&`

* logical *or*: `|`

* logical *equal*: `==`

* logical *not equal*: `!=`

* logical *greater*: `>`

* logical *less*: `<`

* logical *less than or equal to*: `<=`

* logical *greater than or equal to*: `>=`

### arrange()
To order data sets, the `arrange()` function can be used. You can sort either in ascending or descending order. For example, by working hours:

```{r arrange1, eval =TRUE}
pssAsc <- arrange(
  pss, 
  wkhtot
)
head(pssAsc)
```

Using the `desc()` function within `arrange()`, cases are sorted in descending order. `desc` stands for *descending*. Alternatively, you can simply put a minus sign before the variable name to get cases sorted in descending order as well.

```{r arrange2, eval = TRUE}
pssDesc <- arrange(
  pss,
  desc(wkhtot)
)

head(pssDesc)

# Alternativ Minuszeichen vor Variable
pssDesc2 <- arrange(
  pss, 
  -wkhtot
)
```

### mutate()
To calculate new variables or recode a variable, `mutate()` is used. In this example, we calculate a variable that outputs the difference from the average working hours within our survey.
``` {r mutate1, eval=FALSE}
mutate(
  pss, 
  wkhtotCen = wkhtot - mean(wkhtot, na.rm = TRUE)
)
```
```{r mutate1-table, eval=TRUE, echo=FALSE}
paged_table(
  mutate(
    pss, 
    wkhtotCen = wkhtot - mean(wkhtot, na.rm = TRUE)
  ),
  options = list(rows.print = 15)
)
```

If we want to create new dummy or categorical variables, we need to use the `case_when()` function additionally. In the `case_when()` function, you hierarchically specify which condition to check and how to recode. For example, we could recode the variable `district` and turn the *character* variable into an *integer* variable. It is important that the checks are done hierarchically, similar to an **if statement**, from the most specific condition to the most general condition (otherwise, there will be nonsensical variables!).

```{r mutate3, eval=FALSE}
mutate(
  pss,
  districtRec = case_when(
    district == "Distrikt 1" ~ 1, 
    district == "Distrikt 5" ~ 5, 
    district == "Distrikt 7" ~ 7,
    district == "Distrikt 10" ~ 10, 
    district == "Distrikt 12" ~ 12,
  )
)
```
```{r mutate3-table, eval=TRUE, echo=FALSE}
paged_table(
  mutate(
    pss,
    districtRec = case_when(
      district == "Distrikt 1" ~ 1, 
      district == "Distrikt 5" ~ 5, 
      district == "Distrikt 7" ~ 7,
      district == "Distrikt 10" ~ 10, 
      district == "Distrikt 12" ~ 12,
    )
  ),
  options = list(rows.print = 15)
)
```

Multiple conditions can also be linked here: Now we want to calculate a dummy variable that indicates whether people live in District 12 and are female (*female*).

```{r mutate4, eval=FALSE}
mutate(
  pss, 
  d12gndr = case_when(
    district == "Distrikt 12" & gndr == "female" ~ 1
  )
)

``` 
```{r mutate4-table, echo=FALSE, eval=TRUE}
paged_table(
  mutate(
    pss, 
    d12gndr = case_when(
      district == "Distrikt 12" & gndr == "female" ~ 1
    )
  ),
  options = list(rows.print = 15)
)
```

In this example, we have only specified one condition for the value `1` on the new variable. As we can see, all other cases are automatically assigned `NA`. However, we want to assign the value `0` to all other cases. To avoid having to write out the different combinations, we use another argument of the `case_when()` function:

```{r mutate5, eval=FALSE}
mutate(
  pss, 
  d12gndr = case_when(
    district == "Distrikt 12" & gndr == "female" ~ 1,
    TRUE ~ 0
  )
)
```
```{r mutate5-table, echo=FALSE, eval=TRUE}
paged_table(
  mutate(
    pss, 
    d12gndr = case_when(
      district == "Distrikt 12" & gndr == "female" ~ 1,
      TRUE ~ 0
    )
  ),
  options = list(rows.print = 15)
)
```

With the argument `TRUE ~ 0`, we specify that all other values will be assigned the value `0`. This way, we do not need to write out all other combinations as code.

### summarize()
The `summarize()` function allows us to summarize a value from columns. For example, the mean of a column.
``` {r summarize, eval = TRUE}
summarize(
  pss, 
  mean(wkhtot)
)
```

Any function that requires a column as input can be used: including `first()`, `last()`, `nth()`, `n()`, `n_distinct()`, `IQR()`, `min()`, `max()`, `mean()`, `median()`, `var()`, and `sd()`.

### group_by()
Often in datasets, we have categorical variables by which we want to group the dataset. For example, we could group the dataset by field of study. We do this with the `group_by()` function:

```{r group-by, eval = FALSE}
group_by(
  pss,
  gndr
)

```

Apparently, nothing has changed in the dataset, but here is the important difference from `arrange()`: `group_by()` does not sort the dataset, but rather groups the dataset. Therefore, the output does not change. If, for example, we want to display the mean semester by field of study, we concatenate `group_by()` and `summarize()` (we will do this more elegantly below with the **Piping Operator**):
```{r sum-groups, eval = TRUE}
summarize(
  group_by(
    pss, 
    gndr
  ), 
  mean(wkhtot)
)

```

Here we can see what `group_by()` does: Instead of one mean, we now get four means here (one for each course of study in the dataset). **Important**: Grouping should always be resolved subsequently with the function `ungroup()`.
