---
title: "2 Combining Datasets"  # Page Title
weight: 315  # Custom Weight 
menuTitle: "" # Short Title if the title is too long
tags: ["dplyr", "tidyverse", "dataset"]  # Insert tags here; Abbreviation of what happens on the page
---
```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
library("rmarkdown")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")
pssMacro <- readRDS("./data/pssMacro.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Often, we have micro-level datasets (e.g., European Social Survey) that we want to combine with macro datasets (e.g., country-level economic data). For this purpose, `dplyr` offers a total of four functions, with `left_join()` being used in most cases.

Let's consider the following example: We have found a second dataset during our research that lists the percentage of residents in a district affected by poverty (`poverty`) and the percentage of the population experiencing malnutrition (`nutrition`) in addition to our **PSS** dataset.

```{r, echo=FALSE, eval=TRUE}
kable(
  pssMacro,
  caption =  "Makrodaten pro Distrikt"
)
```

Suppose we now want to include in a multilevel model how these district factors affect satisfaction with democracy. To do this, the two variables `maxsem` should be added to the `pss` dataset. We will use `left_join()` for this purpose:
```{r merging, eval=TRUE}
pssMerged <- pss %>%
  left_join(
    pssMacro, 
    by = "district"
  )
```
```{r merging-table, eval=TRUE, echo=FALSE}
paged_table(
  pssMerged,
  options = list(
    rows.print = 15,
    cols.print = 16
  )
)
``` 

Also, you could use `right_join()` here:
```{r merging2, eval=TRUE}
pssMerged2 <- pssMacro %>%
  right_join(
    pss, 
    by = "district"
  ) 
```
```{r merging2-table, eval= TRUE, echo=FALSE}
paged_table(
  pssMerged2,
  options = list(
    rows.print = 15,
    cols.print = 16
  )
)
``` 

For example, in multilevel models, we could calculate the effects of individual factors such as field of study and high school grades, as well as the effects of supervision ratio and seminar size.
