---
title: "Merging Datasets (Adding Cases)"  # Page Title
weight: 314  # Custom Weight
menuTitle: "Add Cases"  # Short Title if necessary
tags: ["dplyr", "tidyverse", "dataset"]  # Tags: brief description of page content
---
```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")
pss1 <- readRDS("./data/pss1.rds")
pss5 <- readRDS("./data/pss5.rds")
pss7 <- readRDS("./data/pss7.rds")
pss10 <- readRDS("./data/pss10.rds")
pss12 <- readRDS("./data/pss12.rds")
```

Im nächsten Schritt nehmen wir nun an, dass die Datenerfassung getrennt nach Distrikt durchgeführt wurde und es somit fünf Teildatensätze gibt, die nun zu einem vollständigen Datensatz verbunden werden sollen. Dazu nutzen wir die Funktion `bind_rows()`. In unserem Beispiel haben alle fünf Teildatensätze genau die gleiche Anzahl an Variablen, die dazu auch noch genau gleich benannt sind! Mit dem Argument `.id` erstellen wir eine Variable namens `"origin"`, die die Herkunft des Falles erfasst. Dies ist automatisch nummeriert. Mit `mutate()` machen wir daraus einen Faktor, der eine bessere Beschreibung beinhaltet (*process1*, *process2*, *process3*, *process4*, *process5*)

```{r merge1, eval=TRUE}
pssAll <- pss1 %>%
  bind_rows(
    list(
      pss5,
      pss7, 
      pss10, 
      pss12
    ), 
    .id = "origin"
  ) %>%
  mutate(
    origin = factor(
      origin, 
      labels = c(
        "process1", 
        "process2", 
        "process3", 
        "process4",
        "process5"
      )
    )
  )

table(pssAll$origin)
head(pssAll$origin)
``` 

Wir haben hier jetzt also aus fünf Teildatensätzen einen gesamten Datensatz erstellt, der alle Fälle der fünf Teildatensätze enthält. Wichtig, in diesem Fall waren alle Variablennamen gleich! 

Now we will try out what happens, for example, if there is a typo in a subset of data. First, let's create two new datasets, each including only 3 cases, and different variables.

``` {r merge-prob, eval=TRUE}
pssA <- pss[1:3, 2:3]

District <- c(
  "Distrikt 1", 
  "Distrikt 5",
  "Distrikt 7"
)

gndr <- c(
  "male",
  "female",
  "female"
)

pssB <- data.frame(
  District, 
  gndr
)

head(pssA)

head(pssB)
``` 

So, in both datasets, we have two variables indicating District and Gender. However, in dataset `pssB`, the District variable is spelled differently. Let's try `bind_rows()`.

``` {r test-bind-rows, eval=TRUE}
pssTest <- pssA %>% 
  bind_rows(pssB)

pssTest
```

Since the variable names are not exactly the same, three variables are now created: `district`, `gndr`, and `District`. Where the variable is missing, `NAs` are automatically generated. This can be advantageous but can also become tricky if the dataset creation did not strictly follow a coding name schema. **Solution**: Clarify and rename variables beforehand. Otherwise, you can use `full_join()`.

### Merging Datasets (Different Column Names)
This approach is no less complex than renaming column names but still offers an alternative. With `full_join()`, we combine two datasets and can specify in the `by` argument which columns have the same content. The drawback here is that the same column names must also be listed, as otherwise (as in this example) the variables `gndr.x` and `gndr.y` will be created. This is because `full_join()` is actually intended to add new/additional variables.

In our example, we would specify that from dataset `pssA`, the column `district` is equal to the column `District` from dataset `pssB`. The same applies to the `gndr` variable.

```{r full-join, eval=TRUE}
pssTest2 <- pssA %>% 
  full_join(
    pssB,
    by = c(
      "district" = "District", 
      "gndr" = "gndr"
    )
  )

head(pssTest2)
```

This is how you can successfully merge datasets with different variable names!
