---
title: "Piping"  # Page Title
weight: 312 # Custom Weight 
menuTitle: "" # Short Title if the Title is too long
tags: ["dplyr", "tidyverse", "piping"]  # Insert tags here; Short word describing what happens on the page
---
```{r setup, include = FALSE}
library("knitr")
library("htmltools")
library("tidyverse")
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Now that the most common functions have been introduced, we will explain how modularization or **piping** works with `dplyr`. As mentioned above, one immense advantage of `dplyr` is that operations are **piped**, meaning they are divided into individual components that are easy to follow.

Through **piping**, datasets or **tibbles** are passed on from previous operations. This is done using the `%>%` operator. Let's look at a first example:
``` {r pipe1, eval = TRUE}
pss <- pss %>%
  group_by(district) %>%
  mutate(
    wkhtotMean = mean(
      wkhtot, 
      na.rm = TRUE
    )
  ) %>%
  ungroup() 
```
First, we pass the loaded dataset `pss`. Then we perform three operations on the dataset: `group_by()`, `mutate()`, and `ungroup()`. Afterwards, we pass the modified object back to the original object using the assignment arrows `->` (so we are overwriting it!).

This example calculates the average working hours by district and stores it in the new variable `wkhtotMean`. To ensure this is also saved in the dataset `pss`, we save these steps back into the object `pss` using the assignment arrow.

Since we are *passing* the dataset, we do not need to call it in each operation. The piping operators do not always have to be entered manually; they can be inserted automatically with `[Ctrl] + [Shift] + [M]` (Windows) or `[Cmd] + [Shift] = [M]` (Mac).

Next, we want to perform various preparation steps using **piping**.

### Calculating and Recoding Variables
In this example, we want to create a new variable that distinguishes the level of trust individuals have in politicians (`trstplt`). We want to differentiate between low, medium, and high trust.

```{r pipe2, eval = TRUE}
pss <- pss %>%
  mutate(
    trstpltG = case_when(
      trstplt <= 3 ~ "low", 
      trstplt > 3 & trstplt <= 6 ~ "medium", 
      trstplt > 6 ~ "high"
    )
  )
table(pss$trstpltG)
```

Checking the new variable reveals that it is stored as a *character* type.
```{r trstplt-str, eval = TRUE}
str(pss$trstpltG)
```

This can be changed normally, or it can be written directly into the **piping**:
```{r pipe3, eval = TRUE}
pss <- pss %>%
  mutate(
    trstpltG = case_when(
      trstplt <= 3 ~ "low", 
      trstplt > 3 & trstplt <= 6 ~ "medium", 
      trstplt > 6 ~ "high"
    )
  ) %>% 
  mutate(trstpltG = factor(trstpltG)) #Schritt um von Character auf Factor zu kommen!

table(pss$trstpltG)

str(pss$trstpltG)
```

Now we have a factor, but without order, which we can also fix:
```{r pipe4, eval = TRUE}
pss <- pss %>%
  mutate(
    trstpltG = case_when(
      trstplt <= 3 ~ "low", 
      trstplt > 3 & trstplt <= 6 ~ "medium", 
      trstplt > 6 ~ "high"
    )
  ) %>% 
  mutate(
    trstpltG = factor(
      trstpltG,
      ordered = TRUE, 
      levels = c(
        "low", 
        "medium", 
        "high"
      )
    ) 
  )

table(pss$trstpltG)

str(pss$trstpltG)
```

A more complex task would be: We not only want to calculate the average working hours per district but also the deviation of an individual in the respective district from the district's average!

```{r pipe5, eval = TRUE}
pss <- pss %>%
  group_by(district) %>%
  mutate(
    wkhtotMean = mean(
      wkhtot, 
      na.rm = TRUE
    ),
    wkhtotDist = wkhtot - wkhtotMean  # wir fügen einfach diese einfache Berechnung des Abstands hinzu
  ) %>%
  ungroup()

head(
  pss[, 
      c(
        "district", 
        "wkhtot",
        "wkhtotMean",
        "wkhtotDist"
      )
  ]
)
```

Here we can also see an advantage of the modular principle. When we calculate and *pass on* new variables (here `wkhtotMean`), we can directly use them in the subsequent operations.

Alternatively, we could hierarchically group the data by district and education (`edu`) and then simply output the different means using `summarize()`:

```{r meansdistrictedu, eval = TRUE}
meansDistriktEdu <- pss %>%
  group_by(
    district,
    edu
  ) %>%
  summarize(mean(wkhtot)) 

meansDistriktEdu
```

Let's take a step back and see how we can divide datasets. This is relevant, especially when working with secondary datasets, as they are sometimes collected for more individuals than needed for one's research purposes.

Please provide the Markdown content you would like me to translate.
