---
title: "How to Handle a Data Frame"  # Page Title
weight: 113  # Custom Weight 
menuTitle: "Data Frame in R" # Short Title if the Title is too long
tags: ["dataset", "data frame"]  # Tags: Brief description of what happens on the page
---

```{r setup, include = FALSE}
library("knitr")

knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")   
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Datasets are simply put, a collection of different variables in a data table. In R, this means storing different vectors of the same length in a data table within a new object.

{{% expand \"Why do the vectors in a data table all have the same length?\" %}}
All respondents receive the same questions. Even if specific question groups were only presented to a subset of respondents, the other respondents also have a value on that variable. Namely, an `NA` value. So, they then have a *missing value* on that variable. This also happens when individuals refuse to answer specific questions.
{{% /expand %}}

To get an overview of a dataset, use the `str()` function. `str` stands for *structure*.

```{r structure, eval=TRUE}
str(pss)
```

Here, you can see how many dimensions the data table (**data frame**) has (5000 rows and 14 columns) and you get the value type and the first observations for each variable.

Another way to gain insight into a dataset is by using the `head()` or `tail()` function. These functions display the first six or last six cases, respectively.

```{r head, eval=TRUE}
head(pss)
```

```{r tail, eval=TRUE}
tail(pss)
```

When applying statistical calculations, you do not work with the entire object but usually with individual variables. But how can you address these individual variables? For example, you may want to know the variance of the variable `stfdem`.

We have the **data frame** object named `pss`, and within this object, we have 14 vectors, one of which is named `stfdem`. When you expand the object in the **Environment**, you will see that each variable name is preceded by a `$`. This indicates that the variables are components of the **data frame** and should be addressed as such.

Let's access the component `stfdem`. When accessing objects, we need to call their names:

```{r gettingvar, eval=FALSE}
stfdem
```

{{% expand "Oops! Why is there an error message here?" %}}
The error message indicates that the object `stfdem` does not exist. But it's in the Environment, right?!?! Correct, but as a component of the object `pss`, which needs to be considered when calling it.
{{% /expand %}}

Let's try again:
```{r gettingvar2, eval=TRUE}
pss$stfdem
```

Now it works! So, when you want to address individual variables from a dataset, you need to consider where they are stored in the input. Using the `$` sign indicates that the variable is part of the dataset before the `$` sign.

You now understand the basic structure of a dataset. Next, we will import data into R to start the data analysis. In the following chapter, you will output descriptive statistics of individual variables.
