---
title: "Layout Questions"  # Title of the page
weight: 201  # Custom weight
menuTitle: "Layout" # Short title if necessary
tags: ["ggplot", "layout", "advanced"]  # Tags: brief description of the content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.ub.uni-giessen.de\/methoden-politik\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Within a `ggplot`, almost all displayed components can be modified and adjusted. Some of these modifications will be discussed as follows.

To start, let's create a `ggplot` object with our scatter plot from the introduction to `ggplot`:
``` {r base-scatter, eval=TRUE}
scatter <- ggplot(
  pss, 
  aes(
    stfeco, 
    stfdem
  )
) +  
  geom_jitter(
    alpha = .2, 
    col = "blue"
  ) +
  scale_x_continuous(
    breaks = seq(
      0, 
      10, 
      1
    )
  ) +
  scale_y_continuous(
    breaks = seq(
      0, 
      10, 
      1
    )
  )

scatter
``` 

First, we will add titles, axis labels, and sources again.
``` {r legends, eval=TRUE}
scatterLeg <- scatter +
  labs(
       x = "Satisfaction with Economy",
       y = "Satisfaction with Democracy",
       title = "Correlation Plot",
       caption = "Data: Panem Social Survey.\n Data jittered."
       )

scatterLeg
```

Within the `theme()` function, we can address and modify specific components of the plot. This includes the following properties of the plot:

- `plot.title`
- `axis.title.x` / `axis.title.y`
- `axis.text.x` / `axis.text.y`
- `panel.grid` / `panel.grid.minor` / `panel.grid.major`
- `plot.background` / `panel.background`

{{% notice tip %}}
A complete overview of all settings that can be used in `theme()` can be found in the
[User Documentation](https://ggplot2.tidyverse.org/reference/theme.html).
{{% /notice %}}

We will now make changes gradually. First, we will change the font size, position, and appearance of the title. We do this using `plot.title` in `theme()`. We use the `element_text()` function for this:
``` {r title, eval=TRUE}
scatterLeg +
  theme(
    plot.title = element_text(
      size = 25,
      face = "italic",
      hjust = 0.5
    )
  )
```

For this, we have used the three arguments `size` (font size), `face` (appearance), and `hjust` (position).

Next, we want to edit the axis titles.

``` {r axisticks, eval=TRUE}
scatterAxes <- scatterLeg +
  theme(
    plot.title = element_text(
      size = 25,
      face = "italic",
      hjust = 0.5
    ),
    axis.title.x = element_text(
      size = 16, 
      color = "seagreen", 
      hjust = 0
    ),
    axis.title.y = element_text(
      size = 8, 
      color = rgb(
        0, 
        105, 
        179, 
        maxColorValue = 255
      ), 
      hjust = 1, 
      face = "bold"
    )
  )

scatterAxes
```

Instead of specifying a color, you can determine the hue using the `rgb()` function. Alternatively, you can also use the HTML color code within the `color` argument.

``` {r axisticks2, eval=TRUE}
scatterLeg +
  theme(
    plot.title = element_text(
      size = 25,
      face = "italic",
      hjust = 0.5
    ),
    axis.title.x = element_text(
      size = 16, 
      color = "seagreen", 
      hjust = 0
    ),
    axis.title.y = element_text(
      size = 8, 
      color = "#0069B3", 
      hjust = 1, 
      face = "bold"
    )
  )
```

Now, let's continue experimenting and edit the axis ticks. For this, we use `axis.ticks.x` or `axis.ticks.y`.

```{r axesticks, eval=TRUE}
scatterTicks <- scatterAxes +
  theme(
    axis.text.x = element_text(
      size = 12, 
      angle = 45,
      color = "darkgrey"
    ),
    axis.text.y = element_text(
      size = 11, 
      hjust = 0,
      vjust = 1
    )
  )

scatterTicks
```

With the `angle` argument, we can rotate the axis labels. With `hjust` and `vjust`, we can change the starting position of the text.

Next, we want to change the plot grid, specifically the lines. To do this, we first use the `panel.grid` argument and the `element_line()` function within the argument.
```{r grid1, eval=TRUE}
scatterGrid <- scatterTicks +
  theme(
    panel.grid = element_line(
      color = "green",
      size = 1,
      linetype = "solid" # blank, solid, dashed, dotted, dotdash, longdash, twodash 
    )
  )

scatterGrid
```

The `panel.grid.major` and `panel.grid.minor` arguments can be used to edit the major and minor gridlines separately. For example, if we only want the major gridlines, we do the following:
```{r grid2, eval=TRUE}
scatterGrid <- scatterTicks +
  theme(
    panel.grid.major = element_line(
      color = "green",
      size = 1,
      linetype = "solid" # blank, solid, dashed, dotted, dotdash, longdash, twodash 
    ),
    panel.grid.minor = element_blank()
  )

scatterGrid
```

You can also edit the minor gridlines separately by adding `.x` or `.y` as needed.

Finally, you can also change the background of the plot or panel. This is done using the `plot.background` or `panel.background` arguments. You can use the `element_rect()` function within the argument.

```{r background, eval=TRUE}
scatterGrid +
  theme(
    plot.background = element_rect(
      color ="darkgray",
      size = 2,
      fill = "lightpink"
    ),
    panel.background = element_rect(
      fill = "moccasin"
    )
  )
```

There are also a variety of pre-made themes that can be further customized.

{{% notice tip %}}
An overview of available themes can be found [here](https://ggplot2.tidyverse.org/reference/ggtheme.html).
{{% /notice %}}
