---
title: "naniar & UpSetR"  # Page Title
weight: 204  # Custom Weight 
menuTitle: "naniar & UpSetR" # Short Title if the Title is too long
tags: ["ggplot", "Missing values", "advanced"]  # Insert tags here; Abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("naniar")
library("UpSetR")
pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.ub.uni-giessen.de\/methoden-politik\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Using the `naniar` package makes the steps shown above much faster and easier to display. The package also always creates a `ggplot`, so the adjustments learned above are also possible here. First, we use functions to display tables using `naniar`. The first one is the `miss_var_summary()` function, which gives us the absolute and relative frequency of *missing values* in the variables.

```{r table-naniar1, eval=TRUE}
pss %>%
  miss_var_summary()
```

We can also group this:
```{r table-naniar2, eval=TRUE}
pss %>% 
  group_by(district) %>% 
  miss_var_summary()
```

First, we can display a distribution of *missing values* in the dataset. The `gg_miss_var_cumsum()` function gives us the cumulative sum of *missing values* per variable. This way, we can see how the *missing values* are distributed across the variables.```

```{r naniar0, eval=TRUE}
gg_miss_var_cumsum(pss)
```

The function `vis_miss()` visualizes the *missing values* of an entire dataset (unless we specify a subset).

```{r naniar1, eval=TRUE}
vis_miss(pss)
```

Another appealing alternative is the function `gg_miss_upset()` from the package `naniar`. Here, the frequencies of the combinations of *missing values* between variables are also displayed. However, this can quickly become overwhelming with very large datasets. It can be insightful for subsets (e.g., when checking if individuals have only partially answered an item battery or have not answered the item battery at all).

```{r naniar2, eval=TRUE}
gg_miss_upset(pss)
```

In the graph, it can be seen that the four variables `trstprt`, `trtprl`, `stfdem`, and `agea` have missing values. The following combinations exist:

- 311 cases with missing values in `edu`,
- 148 cases with missing values in `agea`,
- 82 cases with missing values in `stfdem`,
- 30 cases with missing values in `trstprl`,
- 13 cases with missing values in `trstprt`,
- 9 cases with missing values in both `stfdem` and `edu`,
- 5 cases with missing values in both `agea` and `edu`,
- 4 cases with missing values in both `trstprl` and `edu`,
- 3 cases with missing values in both `stfdem` and `agea`,
- 3 cases with missing values in both `trstprt` and `edu`,
- 1 case with missing values in both `trstprl` and `stfdem`,
- 1 case with missing values in both `trstprt` and `agea`.

Overall, the maximum number of combinations is calculated as follows: $2^{Number of Variables} - 1$. In this case, there would be 31 possible combinations, but only 12 are displayed. Why?

Additionally, *missing values* of two variables can be easily visualized in a `ggplot` using the function `geom_miss_point()`:
```{r missValggplot, eval=TRUE}
ggplot(
  pss,
  aes(
    x = district,
    y = agea
  )
) +
  geom_miss_point() 
```

This allows for easy identification of any potential clustering of *missing values* in specific combinations.

Alternatively, the functions `gg_miss_var()` and `gg_miss_fct()` can also be used.

The function `gg_miss_var()` displays the number of *missing values*. Using the `facet` argument, this can be broken down by individual categories. This helps in identifying if one group has significantly more *missing values* than another group.
```{r ggmissvar, eval=TRUE}
gg_miss_var(
  pss,
  facet = district
)

```

The function `gg_miss_fct()` provides a visually appealing representation of *missing values*.

``` {r ggmissfct, eval=TRUE}
gg_miss_fct(
  x = pss, 
  fct = district
)
```
You can also display this based on values of another variable to see if there are strong group differences:
``` {r ggmissfct2, eval=TRUE}
gg_miss_fct(
  x = pss, 
  fct = district
) + 
  labs(title = "NA in PSS nach Distrikt")
```

That's it for *missing values*!
