---
title: "Missing values"  # Page title
weight: 203  # Custom weight 
menuTitle: "Missing values" # Short title if the title is too long
tags: ["ggplot", "Missing values", "advanced"]  # Tags; brief description of the page content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("naniar")
library("UpSetR")
pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.ub.uni-giessen.de\/methoden-politik\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Often, before starting the actual data analysis, it is important to inspect the data and especially check for ***missing values***. There are two comprehensive packages that build on `ggplot2` for this purpose: `naniar` and `UpSetR`.

First, we need to install or load the packages:
``` {r packages-install-mis, eval=FALSE}
install.packages("UpSetR")
install.packages("naniar")

library("UpSetR")
library("naniar")
```

Next, we will continue using the dataset `pss`. Our goal is to visualize the *missing values* per variable. To do this, we first filter the dataset for the ID variable and the following four variables: `stfdem`, `stfeco`, `trstprl`, and `trstlgl`. Then, we convert the dataset into a long format and create a third column indicating whether it is a *missing* value or not. We then group by variables and the new third column, count the *missing values* per variable (as well as the non-missing values), exclude the non-missing values, and sort the table in descending order. This will show us the number of *missing values* in each of the four variables.
``` {r miss-tidy, eval=TRUE}
missingValues <- pss %>%
  select(
    c(
      "idno", 
      "stfdem",
      "stfeco", 
      "trstprl", 
      "trstlgl"
    )
  ) %>% 
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "val"
  ) %>%
  mutate(is.missing = is.na(val)) %>%
  group_by(
    variable, 
    is.missing
  ) %>%
  summarize(
    num.missing = n()
  ) %>%
  filter(is.missing == TRUE) %>%
  select(-is.missing) %>%
  arrange(desc(num.missing))

missingValues
```

Next, you can display a simple bar plot with this new dataset:
```{r missValBarplot, eval=TRUE}
missingValues %>%
  ggplot() +
  geom_bar(
    aes(
      variable, 
      num.missing
    ), 
    stat = 'identity'
  ) +
  labs(
    x = 'Variable',
    y = "Anzahl MV", 
    title = 'Missing Values pro Variable'
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45, 
      hjust = 1
    )
  )
```

Here you can test all the tricks from above. However, we now want to display percentages to see how much percent of the variable *missings* there are. To do this, we repeat the steps from before, but add a `mutate` step that gives us the percentages.

```{r missValPercent, eval=TRUE}
#Prozente
missingValues <-  pss %>%
  select(
    c(
      "idno", 
      "stfdem",
      "stfeco", 
      "trstprl", 
      "trstlgl"
    )
  ) %>% 
  pivot_longer(
               everything(),
               names_to = "key",
               values_to = "val"
               ) %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(
           key, 
           total,
           isna
           ) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)

levels <- (
           missingValues  %>% 
             filter(isna == T) %>%     
             arrange(desc(pct))
           )$key

missingsPercent <- missingValues %>%
  ggplot() +
  geom_bar(
           aes(
               x = reorder(
                           key, 
                           desc(pct)
                           ), 
               y = pct, 
               fill = isna
               ), 
           stat = 'identity',
           alpha = 0.8) +
  scale_x_discrete(limits = levels) +
  scale_fill_manual(
                    name = "", 
                    values = c(
                               'steelblue', 
                               'tomato3'
                               ), 
                    labels = c(
                               "vorhanden",
                               "fehlend"
                               )
                    ) +
  coord_flip() +
  labs(
       title = "Prozent von missing values", 
       x = 'Variable', 
       y = "% missing values"
       )

missingsPercent
```

Alternatively, you can also display the *missings* in a way that shows which case is missing on which variable. However, this quickly becomes confusing with large datasets.
``` {r missValCasewise, eval=FALSE}
# pro Fall (wird aber bei großen Datensätzen etwas schwer zu lesen)
rawPlot <-  pss %>%
  select(
    c(
      "idno", 
      "stfdem",
      "stfeco", 
      "trstprl", 
      "trstlgl"
    )
  ) %>% 
  pivot_longer(
               -c("idno"),
               names_to = "key",
               values_to = "val"
               ) %>%
  mutate(isna = is.na(val)) %>%
  ggplot(
         aes(
             key, 
             idno, 
             fill = isna)
        ) +
  geom_raster(alpha = 0.8) +
  scale_fill_manual(
                    name = "",
                    values = c(
                               "steelblue", 
                               "tomato3"
                               ),
                    labels = c(
                               "vorhanden",
                               "fehlend"
                               )
                    ) +
  scale_x_discrete(limits = levels) +
  labs(
       x = "Variable",
       y = "Row Number",
       title = "Missing values in rows"
       ) +
  coord_flip()

rawPlot
```

Let's move on to the **libraries** `naniar` and `UpSetR`!
