---
title: "ggboxplot()"  # Page Title
weight: 132  # Custom Weight 
menuTitle: "ggboxplot()" # Short Title if the Title is too long
tags: ["`ggboxplot()`", "ggplot"]  # Tags; Brief description of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("ggpubr")

cbp1 <- c(
  "#999999", 
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442", 
  "#0072B2",
  "#D55E00",
  "#CC79A7"
)
pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Alternatively, you can use the `ggboxplot()` function from the **library** `ggpubr`: First, install the **library** and then activate it. Now, we will display working hours by gender.
```{r bp1, eval=TRUE}
ggbox <- ggboxplot(
  pss,
  x = "gndr",              
  y = "wkhtot",             
  color = "ivory4",
  fill = "gndr",
  palette = c(
    "lightblue", 
    "pink"
  ),
  ylab = "Total work hours (incl. overtime)",
  xlab = "Gender (1:male, 2:female)",
  na.rm = TRUE
) 

ggbox
```

Here, we can make fine adjustments easily: For example, we can add a horizontal line (`geom_hline()`) to mark the mean for each group. 
```{r bp2, eval=TRUE}
ggbox +
  geom_hline(
    aes(
      yintercept = mean(
        wkhtot[gndr == "male"], 
        na.rm = TRUE
      )
    ),
    lty = 2,
    lwd = 1,
    color = "lightblue") + 
  geom_hline(
    aes(
      yintercept = mean(
        wkhtot[gndr == "female"], 
        na.rm = TRUE
      )
    ),
    lty = 2,
    lwd = 1, 
    color = "pink"
  ) +
  theme(legend.position = "none")
```

Copy the code into your script and try different numerical values for the `lwd` and `lty` arguments.

{{% expand \"What do the arguments lwd and lty stand for?\" %}}
`lwd` stands for *line width*. This specifies the thickness of the line. `lty` stands for *linetype*, allowing you to choose different line types. You can find an overview [here](http://www.cookbook-r.com/Graphs/Shapes_and_line_types/).
{{% /expand %}}

Next, you can explore `rainclouds` as a better visualization option! However, this is **optional**!
