---
title: "geom_boxplot()"  # Page Title
weight: 131  # Custom Weight 
menuTitle: "geom_boxplot()" # Short Title if the Title is too long
tags: ["geom_boxplot", "ggplot"]  # Tags; Short description of the content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")

cbp1 <- c(
  "#999999", 
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442", 
  "#0072B2",
  "#D55E00",
  "#CC79A7"
)

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

You are familiar with boxplots from your statistics lecture. Boxplots show the quartiles, as well as the minimum and maximum values (or outliers). Now you will learn more about the individual functions on how to create a boxplot. We introduce two functions: `geom_boxplot()` or `ggboxplot()`. We use boxplots to represent a metric variable. Boxplots of a metric variable can also be divided by a grouping variable.

Additionally, on the last page, we introduce **rainclouds**. These are similar to a boxplot but provide additional information on how the values are distributed.

## Boxplots with `geom_boxplot()`
We can easily create boxplots of age. For this, we use the function `geom_boxplot()`:
```{r boxplot, eval=TRUE}
boxplot <- ggplot(
  pss, 
  aes(agea)
) +
  geom_boxplot()

boxplot
```
Now, the x-axis represents the metric variable age. The interpretation of the y-axis may be confusing as it is not interpreted in content. However, `ggplot` requires it for visualization. Therefore, you can simply adjust it and rotate the boxplot:
```{r boxplot2, eval=TRUE}
boxplot <- ggplot(
  pss, 
  aes(agea)
) +
  geom_boxplot() +
  coord_flip() +
  scale_y_continuous(breaks = NULL)

boxplot
```

In the following, we will display the boxplot of age depending on the district. Often, you want to display the distribution of a metric variable based on a categorical variable. To do this, simply add the group variable within the `ggplot()` function in the `aes` argument. **Remember**: We have swapped the axes, so the group variable is on the x-axis (1st argument in `aes`), even though it appears on the y-axis in the plot!
```{r boxplot-district, eval=TRUE}
boxplotDistrict <- ggplot(
  pss, 
  aes(
    district, 
    agea, 
    fill = district
  )
) +
  geom_boxplot() 

boxplotDistrict
```

As before, we can easily adjust the plot:
```{r boxplot-adjustments, eval=TRUE}
boxplotDistrict +
  scale_fill_manual(
                    name = "Distrikt", 
                    values = cbp1
                    ) +
  scale_x_discrete(
                   limits = c(
                              "Distrikt 10", 
                              "Distrikt 7", 
                              "Distrikt 12", 
                              "Distrikt 5",
                              "Distrikt 1"
                              )
                   ) + 
  scale_y_continuous(
    breaks = seq(
      0, 
      100, 
      5
    )
  ) + 
  labs(
    x = "District", 
    y = "Age in years",
    title = "Boxplots of Age by District"
  ) +
  coord_flip()
```

On the next page, you will learn about an alternative!
