---
title: "Assumption 3: Homoscedasticity"  # Title of the page
weight: 304  # Custom weight 
menuTitle: "Homoscedasticity" # Short title if the title is too long
tags: ["regression diagnostics", "homoscedasticity"]  # Insert tags here; brief description of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("lmtest")

pss <- readRDS("./data/pss.rds")

olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)      
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

The third assumption is **Homoscedasticity**. **Homoscedasticity** means that the variance of the residuals is approximately equal and does not vary.

To diagnose whether the variances of the residuals exhibit equal variances, we change the second argument in the `plot()` function to the value `3`:
```{r assumption3, eval = TRUE}
plot(
  olsModel2, 
  3
)
```

$\Rightarrow$ the points should be evenly distributed around the horizontal line. Interpreting this with pseudo-metric data can be challenging.

Using the **library** `ggplot` provides a better visualization as we can jitter the data points. You will learn how to precisely apply `ggplot` in Learning Block 5!

```{r assumption3b, eval = TRUE}
ggplot(
  olsModel2, 
  aes(.fitted, 
      sqrt(abs(.stdresid))
  )
) + 
  geom_point(na.rm = TRUE) +
  geom_jitter(
    width = 0.35,
    height = 0.35
  ) +
  stat_smooth(
    method = "loess",
    na.rm = TRUE,
    se = FALSE
  ) + 
  xlab("Fitted Values") + 
  ylab(expression(sqrt("Standardized residuals"))) + 
  ggtitle("Scale-Location")
```

$\Rightarrow$ Points should be evenly scattered around the horizontal line.

We can also conduct a statistical test. This is helpful when graphical analysis is difficult. This test is called the **Breusch-Pagan test**. 

- $H_0:$ Homoskedastic residuals are present.

We use the `bptest()` function from the **library** `lmtest` for this:
```{r bp-test, eval = TRUE}
library("lmtest")

bptest(olsModel2)
```

The p-value indicates that the null hypothesis cannot be rejected. Therefore, we can assume homoskedastic residuals and consider the assumption fulfilled!

## Help with heteroskedastic data
When dealing with heteroskedastic data, we can perform so-called *robust* estimations. This includes mathematical procedures that are less influenced by heteroskedastic data. The interpretation of the metrics remains exactly the same. To do this, you will use the **library** `robustbase`:
```{r robustlm, eval=FALSE}
install.packages("robustbase")
library("robustbase")

# Just use function lmrob() as lm()
olsModel2rob <- lmrob(
  stfdem ~ 1 + stfeco + trstlgl,
  data = pss
)

summary(olsModel2rob)
```

The result will be complemented with so-called *robustness weights*, which are not of interest to you at the moment. The top part is interpreted as before!
