---
title: "Assumption 2: Normal Distribution of Residuals"  # Title of the page
weight: 303  # Custom weight 
menuTitle: "Normal Distribution" # Short title if necessary
tags: ["regression diagnostics"]  # Insert tags here; Brief description of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")

olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)      
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

To check if the residuals are normally distributed with an expected value of $E=0$, we call the so-called *Normal Q-Q* plot. To do this, we simply change the second argument in the `plot()` function to the value `2`:

```{r assumption2, eval = TRUE}
plot(
  olsModel2,
  2
)

```

$\Rightarrow$ The points should follow the diagonal line or lie closely to it. Therefore, we can also consider the assumption to be met asymptotically here.

We can also analyze the density function of the residuals:
```{r assumption2b, eval = TRUE}
plot(density(olsModel2$residuals))

curve(
  dnorm(
    x, 
    0, 
    summary(olsModel2)$sigma
  ), 
  add = TRUE,
  col = "green", 
  lwd = 2
)
```

$\Rightarrow$ The distribution should follow the green line. This is somewhat imprecise here since we have pseudo-metric data!
