---
title: "Assumption 1: Linear Relationship"  # Page title
weight: 302  # Custom weight
menuTitle: "Lin. Relationship" # Short title if needed
tags: ["regression diagnostics"]  # Tags describing the content of the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("./data/pss.rds")

olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)      
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

To check the assumptions of a linear regression, you can create a plot of the estimated values (*fitted*) and the residuals. In this plot, a line is drawn. If this line is nearly horizontal, the assumption can be considered fulfilled.

To create the plot, use the `plot()` function, where you specify the model to be tested in the first argument and a `1` in the second argument (selecting a specific plot, here `Residuals vs. Fitted`).

```{r assumption1, eval = TRUE}
plot(
  olsModel2,
  1
)

```

$\Rightarrow$ the plot should show us a nearly horizontal line; then the assumption is met, and we can assume a linear relationship.
