---
title: "Assumption 5: Multicollinearity"  # Page title
weight: 306  # Custom weight
menuTitle: "Multicollinearity" # Short title if the title is too long
tags: ["regression diagnostics", "multicollinearity"]  # Tags describing the content of the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("car")

pss <- readRDS("./data/pss.rds")

olsModel2 <- lm(
  stfdem ~ 1 + stfeco + trstlgl,   
  data = pss
)       
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

**Multicollinearity** means that there are linear dependencies among the independent variables. You can check this using the *variance inflation factor* (**VIF**). If the **VIF** value is above $10$, then the **multicollinearity** is high, and the estimation is biased (some recommend a threshold of $5$).

Formula for **VIF**:
$$VIF_i = \frac{1}{1-R^2_i}$$

To calculate the **VIF**, we use the `vif()` function from the **library** `car`. For each independent variable, you will get the **VIF** value:
```{r car, eval=FALSE}
install.packages("car")
library("car")
```
```{r vif, eval=TRUE}
vif(olsModel2)
```

$\Rightarrow$ We do not have any values here that exceed the recommended threshold. So, everything is still okay!
