---
title: "Functions"  # Title of the page
weight: 403  # Custom weight 
menuTitle: "" # If the title is too long, use a short title here
tags: ["functions"]  # Insert tags here; abbreviation of what happens on the page
---

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

## Writing Functions in R

In R, not only can we use the functions provided by packages, but we can also easily write our own functions. First, we will create a custom function to calculate the mean and then proceed to create a function to calculate the interquartile range.

<br> 

### Syntax
Below, we will write our own functions. The basic structure of the code is as follows: 
```{r function, eval=FALSE}
my_function <- function(arg1, arg2, ..., argn){
  # Instructions
}
```

First, we define a name for the function (`my_function`) and assign a function to it. The arguments passed to the function are enclosed in parentheses `()`, and these arguments are used in the instructions (enclosed in curly braces `{}`). These arguments can be datasets, variables, or individual values, depending on how they are used in the instructions.

### Custom Function: Mean
Now, we aim to calculate the mean of a variable using a custom function. To do this, the dataset `statistics` (or any other dataset with a metric variable) should be loaded into the **environment**. We create a function where we pass an argument: the variable for which the mean should be calculated. In the function, we name this argument `x`. Whenever we use `x` in the instructions, the passed argument will be utilized when calling the `own_mean()` function. 

Next, we calculate the mean in the first statement of the new function. Here, we access the passed argument (`x`). In the second statement, we output this value. Then we can call the self-created function and must specify the one required argument (variable of calculation). The mean is calculated and output.
```{r function2, eval=FALSE}
own_mean <- function(x){
  mean = sum(x) / length(x)
  
  print(mean) 
}

own_mean(statistics$grade)
```

Thus, we have written our first custom function!


### Custom Function: Interquartile Range
Now we want to write a function that outputs the lower and upper quartiles, as well as the interquartile range. We need **three statements** in the function:

1. calculation of the lower quartile,
2. calculation of the upper quartile, and
3. calculation of the interquartile range.

Additionally, we output these three values using the `print()` function.
```{r function3, eval=FALSE}
own_iqr <- function(x){
  uGrenze <-  quantile(x,
                       probs = 0.25,
                       na.rm = TRUE
                       )
  
  oGrenze <- quantile(x,
                      probs = 0.75,
                      na.rm = TRUE
                      )
  
  abstand <- oGrenze - uGrenze
  
  print(paste("Das untere Quartil liegt bei:",
              uGrenze
              )
        )
  
  print(paste("Das obere Quartil liegt bei:",
              oGrenze
              )
        )
  
  print(paste("Der Interquartilsabstand beträgt:",
              abstand
              )
        )
}

own_iqr(statistics$grade)
```

However, if we want to store the calculated values of the function in a new object, we need to use `return` as above. First, we create a new list `results`, so that our goal is to have a list object with the individual values. We name the individual parts of the list as in the function `$lowerBound`, `$upperBound`, and `$range`. The `quantile()` function automatically stores a *named numeric*, but we only want to store the numerical value. Therefore, we index the assignment to our list with `[[1]]`.
```{r function4, eval=FALSE}
own_iqr_return <- function(x){
  uGrenze <-  quantile(x,
                       probs = 0.25,
                       na.rm = TRUE
                       )
  
  oGrenze <- quantile(x,
                      probs = 0.75,
                      na.rm = TRUE
                      )
  
  abstand <- oGrenze - uGrenze
  
  print(paste("Das untere Quartil liegt bei:",
              uGrenze
              )
        )
  
  print(paste("Das obere Quartil liegt bei:",
              oGrenze
              )
        )
  
  print(paste("Der Interquartilsabstand beträgt:",
              abstand
              )
        )
  
  results <- list()
  
  results$uGrenze <- uGrenze[[1]]
  
  results$oGrenze <-oGrenze[[1]]
  
  results$abstand <- abstand[[1]]
  
  return(results) 
}

test <- own_iqr(statistics$grade)

test
```

Writing custom functions in R is easy to implement and is helpful in many steps of data analysis when there are no existing functions for specific applications.

### Functions and Loops
We can also effectively use loops in functions. For example, we could write a function that shows us the deviation from the mean and the value from a variable. For this, we use the example dataset `statistics`.

We will now link the function with a **for** loop and an **if** statement. We will name the function `showcase` and pass an argument (the variable to be used). Then the **for** loop will be initiated. If no value is present, a message will be displayed stating that no value is present. If a value is present, the distance to the *mean* will be calculated and this, along with the achieved value, will be displayed.

```{r function-loop, eval = FALSE}
showcase <- function(x){
  for (i in 1:length(x)) {
    if (is.na(x[i])) {
      print(paste("Für Student", 
                  i, 
                  "liegt kein Wert vor."
                  )
            )
      next
    } else {
      abstand = mean(x, 
                     na.rm = TRUE
                     ) - x[i]
      
      print(paste0("Student ",
                  i, 
                  " hat folgenden Wert: ",
                  x[i], 
                  " (Abweichung zum Mean: ",
                  round(abstand, 
                        2
                        ),
                  ")"
                  )
            )
    }
  }
}
```

Now we can display an example for the variable `grade` and `grade2` from the `statistics` dataset.

```{r, eval=FALSE}
showcase(statistics$grade)
```

```{r, eval=FALSE}
showcase(statistics$grade2)
```
