---
title: "Regressions"  # Page title
weight: 306  # Custom weight
menuTitle: "Regressions"  # Short title if the title is too long
tags: ["ggplot", "regression", "advanced"]  # Tags to describe the content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("sjPlot")
library("beyonce")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.ub.uni-giessen.de\/methoden-politik\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

In this example, the goal is to visually represent the variable `stfdem` in all steps. Finally, you want to calculate a regression model on `stfdem` that includes the effects of `stfeco`, `district`, `gndr`, `trstlgl`, `trstprl`, and `agea`.

First, you will now calculate the model before you can start with the graphical representations:
```{r model, eval=TRUE}
model1 <- lm(
  stfdem ~ 1 + stfeco + district + gndr + trstlgl + trstprl + agea,
  pss
)

summary(model1)
```

In the model, you can see significant effects of `stfeco`, `gndrmale`, and `trstlgl`. We will now visualize these effects. But first, you will repeat the plot of all regression coefficients from Learning Block 4 using `dwplot()`.

{{% expand \"Try it yourself first and then check here!\" %}}
\`\`\`R
dwplot(
  model1,
  # you can change the confidence interval here
  ci = 0.95, 
  # you can change the appearance of the points here (all options from geom_point are possible)
  dot_args = list(
    size = 3.5,
    shape = 18
  ),
  whisker_args = list(
    size = 1.25, 
    lty = "solid"
  )
) +
  geom_vline(
    xintercept = 0,
    lty = "dashed",
    color = beyonce_palette(72)[5] 

```markdown
  ) +
  scale_y_discrete(labels = rev(
    c(
      "Satisfaction \n Economy", 
      "District 5 \n(ref: District 1)",
      "District 7 \n(ref: District 1)",
      "District 10 \n(ref: District 1)",
      "District 12 \n(ref: District 1)",
      "Male \n(ref: Female)",
      "Trust Legal System",
      "Trust Parliament", 
      "Trust Age"
    )
  )
  ) + 
  scale_x_continuous(
    breaks = seq(
      -1, 
      1,
      0.2
    ),
    limits = c(-1, 1)
  ) +
  labs(title = "Linear Regression Models on stfdem",
       caption = "Data: Panem Social Survey."
  ) +
  scale_color_manual(values = beyonce_palette(72)[1]) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank()
  )
\`\`\`
{{% /expand %}}

Often, the goal is not only to present regression models in tables but also to graphically represent specific effects, such as the significant effects. Here we present a method: manually through custom **predictions**.

While there are **libraries** like `ggraphExtra`, they have limited plotting capabilities.

However, once we have a multivariate model, this is no longer directly possible. We need to hold the other effects constant to correctly represent the visualization. But it is still easy to implement once you understand what **holding constant** means.

We want to plot the effect of `stfeco` on `stfdem`. This effect should always be interpreted in conjunction with the other effects. To plot the effect, we hold the other effect(s) constant. For this, we use the `sjPlot` **library** and utilize the `plot_model()` function. You also need to load `ggplot`.

```{r sjplot, eval=FALSE}
install.packages("sjPlot")
library("sjPlot")
library("tidyverse")
```

In the `plot_model()` function, you first call the object of the regression model (here `model1`), specify in the second argument that it is a *marginal plot* containing predicted values (`type = pred`), and then determine in the `terms` argument which effect to plot. The function automatically holds all metric variables constant at their mean and categorical variables at the reference category.
```

```{r marginal1, eval=TRUE}
plot_model(
  model1, 
  type = "pred",
  terms = "stfeco"
)
```

So in the graph, you see the effect of `stfeco` on `stfdem`, while holding constant (mean) `trstlgl`, `trstprl`, and `agea` for female (reference level `gndr`) individuals from District 1 (reference level `district`).

If you want to add the effect for different groups of categorical variables, simply specify the additional variable in `terms`:
```{r marginal2, eval=TRUE}
plot_model(
  model1, 
  type = "pred",
  terms = c(
    "stfeco",
    "gndr"
  )
)
```

Similarly, you could easily add the next categorical variable:
```{r marginal3, eval=TRUE}
plot_model(
  model1, 
  type = "pred",
  terms = c(
    "stfeco",
    "gndr",
    "district"
  )
)
```

You can also limit the output to specific `levels` (values of a variable) by placing the `levels` in `[..]` brackets after the variable name (but within the `"..."`):
```{r marginal4, eval=TRUE}
# levels von district
levels(pss$district)

plot_model(
  model1, 
  type = "pred",
  terms = c(
    "stfeco",
    "gndr",
    "district[Distrikt 1, Distrikt 12]")
)
```

Modifications of a `ggplot` work here as well: We change the axes and adjust the colors!
```{r marginal5, eval=TRUE}
plot_model(
  model1, 
  type = "pred",
  terms = c(
    "stfeco",
    "gndr",
    "district"
  )
) +
  scale_x_continuous(
    breaks = seq(
      0, 
      10, 
      1
    )
  ) +
  scale_y_continuous(
    breaks = seq(
      0, 
      10, 
      1
    )
  ) +
  scale_color_manual(values = beyonce_palette(72))
```

{{% tabs %}}
{{% tab name="Task" %}}
Now, represent the relationship between `trstlgl` and `stfdem`. Each district should have its own line, and the plots should be separated by gender. Try it out yourself before looking at the solution.
{{% /tab %}}
{{% tab name="Code" %}}
```R
plot_model(
  model1, 
  type = "pred",
  terms = c(
    "trstlgl",
    "district",
    "gndr"
  )
) +
  scale_x_continuous(
    breaks = seq(
      0, 
      10, 
      1
    )
  ) +
  scale_y_continuous(
    breaks = seq(
      0, 
      10, 
      1
    )
  ) +
  scale_color_manual(values = beyonce_palette(18))
```
{{% /tab %}}
{{% /tabs %}}

That's it! In this chapter, you have practiced creating graphics for individual methods and now have mastered the first steps with `ggplot`!
