---
title: "ggcorrplot()"  # Title of the page
weight: 304  # Custom weight
menuTitle: "ggcorrplot()" # Short title if the title is too long
tags: ["ggplot", "ggcorrplot", "correlation", "advanced"]  # Insert tags here; abbreviation of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("GGally")
library("beyonce")

pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.ub.uni-giessen.de\/methoden-politik\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Correlation plots can also be created using `ggplot`. For this, you will again use the **library** `GGally`.

```{r install-ggally, eval=FALSE}
library("GGally")
```

Now, you will use the function `ggcorr()`: Non-metric variables are automatically excluded.

```{r ggcorr, eval=TRUE}
ggcorr(pss)
```

In the additional argument `method`, you can specify how to handle `NA's` and which type of correlation to calculate (`pearson`, `spearman`, `kendall`):

```{r ggcorr2, eval=TRUE}
ggcorr(
  pss,
  method = c(
    "pairwise",
    "pearson"
  )
)
```

Additionally, you can display the correlation coefficients using `label = TRUE`:

```{r ggcorr3, eval=TRUE}
ggcorr(
  pss,
  method = c(
    "pairwise",
    "pearson"
  ),
  label = TRUE
)
```

You can set the number of decimal places with the argument `label_round`:
```{r ggcorr4, eval=TRUE}
ggcorr(
  pss,
  method = c(
    "pairwise",
    "pearson"
  ),
  label = TRUE,
  label_round = 2
)
```

In the `geom` argument, you can choose between `tile`, `circle`, `text`, or `blank`. 
```{r ggcorr5, eval=TRUE}
ggcorr(
  pss,
  method = c(
    "pairwise",
    "pearson"
  ),
  label = TRUE,
  label_round = 2,
  geom = "circle"
)
```

Lastly, you can specify three colors in the `palette` argument ($-1$, $0$, $1) to create the color scale: I'm using `beyonce` again here! **Important:** You are not providing the entire palette here, but rather a color from the respective palette, hence an additional number in `[..]` brackets!
```{r ggcorr6, eval=TRUE}
ggcorr(
  pss,
  method = c(
    "pairwise",
    "pearson"
  ),
  label = TRUE,
  label_round = 2,
  geom = "circle",
  low = beyonce_palette(72)[1],
  mid = "white", 
  high = beyonce_palette(72)[2]
)
```

Alternatively, you can also change the `limits`. This can be helpful if your correlation values are not very high and all are very weakly colored. With `limits = FALSE`, the endpoints are automatically set (according to the data!).
```{r ggcorr7, eval=TRUE}
ggcorr(
  pss,
  method = c(
    "pairwise",
    "pearson"
  ),
  label = TRUE,
  label_round = 2,
  geom = "circle",
  low = beyonce_palette(72)[1],
  mid = "white", 
  high = beyonce_palette(72)[2],
  limits = FALSE
)
```

**Important:** In `ggcorr()`, non-significant values cannot be hidden, as the author of these functions (rightly) opposes a focus on the significance level.

For more information on the functionality in `ggcorr()`, you can find it [here](https://ggobi.github.io/ggally/reference/ggcorr.html?q=ggcorr).

Now let's move on to the representation of mean comparisons. 

{ /*examples*/ }
