---
title: "Data Types"
weight: 201
menuTitle: "Data Types"
tags: ["data types", "numeric", "integer", "character", "boolean"]
---

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

We have already learned that we can perform calculations directly in R or store the result directly in an object.

However, we can also directly store values in objects, distinguishing four different data types:

1. numeric

1. integer

1. character

1. logical

# Numeric
This data type includes all possible numerical values. Strictly speaking, **integer** is a subtype of this numerical value. Therefore, we can also store decimal values in R. Let's store a simple decimal number in the new object `mathNum` and then call it.

```{r numeric1, eval=TRUE}
mathNum <- 2.345

mathNum
```

Important for handling in the learning module: You can directly copy the code and paste it into your own script. In the top right corner of each code snippet, there is a copy symbol. Simply click on it and then paste it into your script (`cmd` and `V` or `Ctrl` and `V`).

{{% expand \"What do you notice about entering the decimal number?\" %}}
Decimal numbers use a `.` (period) as the separator, not a `,` (comma). This is because R is based on the English language, where the decimal separator is a `.` (period).
{{% /expand %}}

# Integer
We already know this data type, as we used it in the `math` object. **Integer** are whole number values, like the value `7` in the `math` example.

# Character
Next, we have the data type **character**. These are character strings: they can be meaningful words made up of letters, or combinations of letters and numbers or symbols. For example, let's store an object `hometown` with the place of residence (in my case, Berlin).

```{r character1, eval=TRUE}
hometown <- "Berlin"

hometown
```

{{% expand \"What do you notice as a difference when storing a numeric or integer value compared to storing text of type character?\" %}}
When dealing with text, it needs to be enclosed in `"..."` (quotation marks) in the R language. You can use either double `"..."` or single `'...'` for this purpose. **Important:** Stay consistent!
{{% /expand %}}

# logical
This data type plays an important role in data processing, as we will later use conditions that each have the data type **logical**. There are two possible values for the **logical** data type: `TRUE` and `FALSE`. **Important:** These words must be capitalized! Otherwise, R will not recognize them as **logical**. For example, let's store the value `TRUE` in an object named `truth`.

```{r logical1, eval=TRUE}
truth <- TRUE

truth
```

But can we store more than one value?
