---
title: "Dataframes"  # Page Title
weight: 204  # Custom Weight 
menuTitle: "Dataframes" # Short Title if Title is too long
tags: ["datatypes" , "dataframes"]  # Insert tags here; Abbreviation of what happens on the page
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

pss <- readRDS("../../static/files/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

In addition to vectors and factors, there is another important object type for us, the **data frame**. A **data frame** is simply a combination of multiple vectors (variables) of the same length in a matrix. In the conventional format (*wide format*), the variables are found in the columns and the respondents in the rows.

- Columns: Vectors, factors (variables)
- Rows: Cases (individual observation units, e.g., respondents)

Let's illustrate this with the example of the dataset we will use during the course: **Panem Social Survey** (`pss`). This is a training dataset based on the European Social Survey, but with significantly fewer variables/cases (only 10 cases and 4 variables):
```{r df1, eval=TRUE, echo=FALSE}
knitr::kable(pss[1:10, c("idno", "district", "gndr", "agea")])
```

In this example dataset, we have four variables: `idno`, `district`, `gndr`, and `agea`. These are self-explanatory: `idno` is the unique ID, `district` is the district of the respondent, `gndr` is the gender, and `agea` is the age. Often, variables are not intuitively understandable, so you may need to consult a codebook. Handling larger datasets will be covered in the next learning block. 

Let's go to the final challenge!
