---
title: "Start: Histogram"  # Page title
weight: 121  # Custom weight 
menuTitle: "Start" # Short title in case the title is too long
tags: ["histograms", "ggplot"]  # Insert tags here; Brief description of what happens on the page
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

To represent a single continuous variable, we use **histograms**. This allows us to interpret the distribution of the variable. To create a **histogram**, we use the `geom_hist()` function.
```{r hist, eval=TRUE}
hist <- ggplot(
  pss, 
  aes(agea)
) + 
  geom_histogram()  

hist
```

Now, we can customize the graph as before. Start with the titles:
```{r hist-title, eval=TRUE}
hist <-  ggplot(
  pss, 
  aes(agea)
) + 
  geom_histogram(
    color = "lightgray", 
    fill = "gray"
  ) + 
  labs(
    x = "Age in years", 
    y = "Frequency", 
    title = "Histogram of Age (PSS)"
  )

hist
```

Often, we want to interpret the distribution. Therefore, we also display the density distribution. To do this, we add the `geom_density()` function. Additionally, you need to change the axis in `geom_histogram()` to `..density..`!
```{r hist-density, eval=TRUE}
histDens <- ggplot(
  pss, 
  aes(agea)
) + 
  geom_histogram(
    aes(y = ..density..), 
    color = "lightgray", 
    fill = "gray"
  ) + 
  geom_density(
    alpha = 0.2, 
    fill = "lightblue"
  ) +
  labs(
    x = "Age in years", 
    y = "Density", 
    title = "Histogram of Age (PSS)"
       )

histDens
```

Here, too, we can group! 

