---
title: "Grouping"  # Page title
weight: 122  # Custom weight 
menuTitle: "Grouping" # Short title if necessary
tags: ["grouping", "ggplot"]  # Tags; brief description of page content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("beyonce")
pss <- readRDS("./data/pss.rds")
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}}  

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

Similarly, we can **group** histograms of a variable, just like we did with bar charts earlier. To do this, we use the `facet_grid()` function, which divides the plot to make the graphs easier to interpret. In `facet_grid()`, we simply specify in a formula (`x ~ y`) how the variable should be divided on the x-axis or y-axis. Here, we want to divide by gender on the y-axis.
``` {r hist-group, eval=TRUE}
histGroup <- ggplot(
  pss,
  aes(
    agea, 
    fill = gndr
  )
) +
  geom_histogram(
    aes(y = ..density..),
    alpha = 0.2
  ) + 
  geom_density(alpha = 0.2) +
  facet_grid(
    .~ gndr
  ) + 
  scale_fill_manual(
    name = "Gender", 
    labels = c(
      "Female", 
      "Male"
    ),
    values = beyonce_palette(72)
  ) +
  labs(
    x = "Age in years", 
    y = "Density", 
    title = "Histogram of Age"
  )

histGroup
```

## Adjustments
Next, we will modify the labels above the **histograms** and remove the legend (to avoid duplication!).
```{r hist-group2, eval=TRUE}
histGroup <- ggplot(
  pss,
  aes(
    agea, 
    fill = gndr
  )
) +
  geom_histogram(
    aes(y = ..density..),
    alpha = 0.2, 
    bins = 30, 
    position = "identity"
  ) + 
  geom_density(alpha = 0.2) +
  facet_grid(
    .~ gndr 
  ) + 
  scale_fill_manual(values = beyonce_palette(72)) +
  labs(
    x = "Age in years", 
    y = "Density", 
    title = "Histogram of Age"
  ) + 
  theme(legend.position = "none")

histGroup
```

Alternatively, you can also create a Ridgeline plot. More on that on the next page! 

{ /* examples */ }
