---
title: "Ridgeline Plot"  # Page Title
weight: 123  # Custom Weight 
menuTitle: "Ridgeline" # Short Title if necessary
tags: ["ridgline", "ggplot"]  # Tags: brief description of the content
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,   # Code wird angezeigt
  eval = FALSE,  # Code wird nicht ausgeführt
  message = FALSE, # Messages werden nicht angezeigt
  warnings = FALSE # warnings werden nicht angezeigt
)

library("tidyverse")
library("ggridges")

pss <- readRDS("./data/pss.rds")

cbp1 <- c(
  "#999999", 
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442", 
  "#0072B2",
  "#D55E00",
  "#CC79A7"
)
```

{{% buttonShare href=\"https\:\/\/gitlab.com\/bpkleer\/einstieg-in-statistik\/issues\/new?issue[title]=\" icon=\"fas fa-bug\" %}} {{% /buttonShare %}} 

{{% buttonShare href=\"mailto\:\" icon=\"fas fa-paper-plane\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/t.me\/share\/url?url=\" icon=\"fab fa-telegram\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/api.whatsapp.com\/send?text=\" icon=\"fab fa-whatsapp\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/twitter.com\/share?url=\" icon=\"fab fa-x-twitter\" %}} {{% /buttonShare %}}

{{% buttonShare href=\"https\:\/\/www.facebook.com\/sharer\/sharer.php?u=\" icon=\"fab fa-facebook\" %}} {{% /buttonShare %}}

{{% button href=\"https\:\/\/bmc.link\/bpkleerw\" icon=\"fa-solid fa-beer-mug-empty\" %}} {{% /button %}}

An alternative to histograms are **Ridgeline Plots**, which also display the density distribution. To create these, we use a function from the **library** `ggridges`. **Ridgeline Plots** always represent split variables, so you need to specify a y-variable in the `ggplot()` function. Let's create a **Ridgeline Plot** of age separated by gender categories.

```{r rlplot, eval=FALSE}
install.packages("ggridges")
library("ggridges")
```

To create a **Ridgeline Plot**, we use the function `geom_density_ridges()`: Within the `ggplot()` function, we set gender as the y-variable and specify that the color representation should differ based on the variable `gndr` (argument `fill = gndr`):
```{r rlplot1, eval=TRUE}
ridgeline <- ggplot(
  pss, 
  aes(
    agea,
    gndr,
    fill = gndr
  )
) +
  geom_density_ridges(
    scale = 0.9,
    alpha = 0.4
  ) +
  theme_ridges() +
  theme(legend.position = "none")

ridgeline
```

You can easily customize this as well.
```{r rlplot3, eval=TRUE}
ridgeline + 
  labs(
    title = "Ridgeline Plot", 
    x = "Age in years",
    y = ""
  ) + 
  theme_ridges(
    grid = TRUE,
    center_axis_labels = TRUE
  ) + 
  scale_fill_cyclical(
    labels = c(
      "female", 
      "male"
    ),
    values = cbp1,
    guide = "legend",
    name = "Gender"
  ) +
  theme(axis.text.y = element_blank())
```
